---
title: "Source summary and comparisons"
author: "Emily Moore"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(phsmethods)
library(lubridate)
library(tidyr)
```

```{r load_data, include=FALSE}
extracts_path<- "/PHI_conf/Dementia_Index/data/extracts/"
clean_path <-"/PHI_conf/Dementia_Index/data/cleaned_extracts/"
smr_all<- readRDS(paste0(clean_path , "smr_diags_grouped.rds") )
                  
PDS <- readRDS("/PHI_conf/Dementia_Index/data/cleaned_extracts/PDS_clean.rds") 
dementia_deaths <- readRDS("/PHI_conf/Dementia_Index/data/extracts/dementia_deaths.rds")
CHC <- readRDS("/PHI_conf/Dementia_Index/data/cleaned_extracts/CHC_clean.rds")
social_care <- readRDS("/PHI_conf/Dementia_Index/data/cleaned_extracts/SC_clean.rds")

##this is just the last 2yrs for now
pis <-readRDS( "/PHI_conf/Dementia_Index/data/cleaned_extracts/pis_clean.rds")

```

## Compare CHI vs PDS


Quite a large number of people that do not appear in PDS do appear in other sources.

Care home census has the largest proportion of CHI not found in PDS, even if limited to those with medically diagnosed dementia. 


```{r chi_compare, echo=FALSE}
smr_chi_list <- smr_all %>% 
  mutate(year_diag = year(smr_diagnosis_date_1)) %>%
           group_by(smr_upi_number, year_diag) %>% slice(1) %>% ungroup() %>%
  mutate(valid_chi = chi_check(smr_upi_number))

PDS_chi_list <- PDS %>% group_by(pds_chi_number) %>% slice(1) %>%
  mutate(inpds = 1)%>% select(inpds, pds_chi_number)
##smrs  
smr_chi_list <- smr_chi_list %>% left_join(PDS_chi_list, by=c("smr_upi_number" = "pds_chi_number")) %>%
  mutate(inpds = case_when(is.na(inpds)~0, T~inpds)) 

overall_smr_chi <- smr_chi_list %>% group_by(smr_upi_number) %>% slice(1) %>% 
  ungroup() %>% group_by(inpds) %>%count() %>%
  pivot_wider(names_from = inpds, values_from = n, names_prefix = "CHIinPDS") %>%
  mutate(perc_smr_chi_in_pds = round(CHIinPDS1/(CHIinPDS1+CHIinPDS0)*100,1)) %>% mutate(source = "SMRA")
###deaths
deaths_chi <- dementia_deaths %>% select(upi_number, date_of_death) %>% 
  left_join(PDS_chi_list, by = c("upi_number" = "pds_chi_number")) %>%
  mutate(inpds = case_when(is.na(inpds)~0, T~inpds)) 

deaths_overall_chi <- deaths_chi %>% group_by(inpds) %>% count() %>%
   pivot_wider(names_from = inpds, values_from = n, names_prefix = "CHIinPDS") %>%
  mutate(perc_smr_chi_in_pds = round(CHIinPDS1/(CHIinPDS1+CHIinPDS0)*100,1)) %>%
  mutate(source = "NRS deaths")
##chc
chc_list <- CHC  %>% select(CHC_UPI_NUMBER, CHC_first_admission) %>% 
  left_join(PDS_chi_list, by=c("CHC_UPI_NUMBER" = "pds_chi_number")) %>%
  mutate(inpds = case_when(is.na(inpds)~0, T~inpds)) 
chc_overall <- chc_list %>% 
  group_by(CHC_UPI_NUMBER) %>% slice(1) %>% 
  ungroup() %>% group_by(inpds) %>%count() %>%
  pivot_wider(names_from = inpds, values_from = n, names_prefix = "CHIinPDS") %>%
  mutate(perc_smr_chi_in_pds = round(CHIinPDS1/(CHIinPDS1+CHIinPDS0)*100,1)) %>% mutate(source = "Care Home Census")
###CHC MD only
chc_list_md <- CHC  %>% 
  filter(CHC_dementia_subtype=="07 Yet to be determined") %>%
    select(CHC_UPI_NUMBER, CHC_first_admission) %>% 
  left_join(PDS_chi_list, by=c("CHC_UPI_NUMBER" = "pds_chi_number")) %>%
  mutate(inpds = case_when(is.na(inpds)~0, T~inpds)) 
chc_overall_md <- chc_list_md %>% 
  group_by(CHC_UPI_NUMBER) %>% slice(1) %>% 
  ungroup() %>% group_by(inpds) %>%count() %>%
  pivot_wider(names_from = inpds, values_from = n, names_prefix = "CHIinPDS") %>%
  mutate(perc_smr_chi_in_pds = round(CHIinPDS1/(CHIinPDS1+CHIinPDS0)*100,1)) %>% mutate(source = "Care Home Census - MD only")


###SC

SC_list <- social_care  %>% select(upi_number, diagnosis_date) %>% 
  left_join(PDS_chi_list, by=c("upi_number" = "pds_chi_number")) %>%
  mutate(inpds = case_when(is.na(inpds)~0, T~inpds)) 
SC_list <- SC_list %>% 
  group_by(upi_number) %>% slice(1) %>% 
  ungroup() %>% group_by(inpds) %>%count() %>%
  pivot_wider(names_from = inpds, values_from = n, names_prefix = "CHIinPDS") %>%
  mutate(perc_smr_chi_in_pds = round(CHIinPDS1/(CHIinPDS1+CHIinPDS0)*100,1)) %>% mutate(source = "Social care")

#

##pis
pis_chi <- pis %>% group_by(pis_pat_upi_c) %>% slice(1) %>%
  ungroup() %>%
  left_join(PDS_chi_list, by=c("pis_pat_upi_c" = "pds_chi_number")) %>%
  mutate(inpds = case_when(is.na(inpds)~0, T~inpds)) 
pis_overall <-pis_chi %>% 
  group_by(pis_pat_upi_c) %>% slice(1) %>% 
  ungroup() %>% group_by(inpds) %>%count() %>%
  pivot_wider(names_from = inpds, values_from = n, names_prefix = "CHIinPDS") %>%
  mutate(perc_smr_chi_in_pds = round(CHIinPDS1/(CHIinPDS1+CHIinPDS0)*100,1)) %>% mutate(source = "PIS")
#table(pis_chi$pis_pat_upi_c %in% PDS$pds_chi_number)
##69% in PDS - so quite good but this is recent years only!


#
df <- rbind(chc_overall,chc_overall_md, overall_smr_chi, deaths_overall_chi,SC_list, pis_overall)
df <- df %>% select(source, everything())
```

```{r table1,  echo=FALSE}
knitr::kable(df, caption = "Proportion of CHI in other datasets with dementia diagnosis that are found in PDS", 
             col.names = c("data source", "N CHI in PDS", "N CHI not found in PDS", "% CHI in source also found in PDS"))
```


## Compare CHI vs PDS - Time trends

There are a higher proportion of CHI found not in PDS but found in other sources in the earlier years of the data. However, there is still quite a high proportion missing from PDS in other years. The year is the year the person is first foud in the other data source, the CHI is searched for in all years of PDS, it does not have to be in the same year as the other source record.



### SMR CHI not in PDS by year

For admissions before 2017, a very low proportion of CHI in SMR sources are found in PDS. The proportion increases to 22% in 2017 and steadily increases after that, reaching 60% in 2024. 

```{r chi_compare_yr_smr,  echo=FALSE}
smr_chi_list <- smr_all %>% 
  mutate(year_diag = year(smr_diagnosis_date_1)) %>%
           group_by(smr_upi_number, year_diag) %>% slice(1) %>% ungroup() %>%
  mutate(valid_chi = chi_check(smr_upi_number))

PDS_chi_list <- PDS %>% group_by(pds_chi_number) %>% slice(1) %>%
  mutate(inpds = 1)
  
smr_chi_list <- smr_chi_list %>% left_join(PDS_chi_list, by=c("smr_upi_number" = "pds_chi_number")) %>%
  mutate(inpds = case_when(is.na(inpds)~0, T~inpds)) 

overall_smr_chi <- smr_chi_list %>% group_by(smr_upi_number) %>% slice(1) %>% 
  ungroup() %>% group_by(inpds) %>%count()

annualsmr_chi <- smr_chi_list %>%  group_by(inpds, year_diag) %>% count() %>% 
  pivot_wider(names_from = inpds, values_from = n, names_prefix = "In_PDS")
 annualsmr_chi <- annualsmr_chi %>%
   mutate(year_diag = case_when(year_diag < 2014 ~"before 2014", T~ as.character(year_diag))) %>% 
   group_by(year_diag) %>% 
   summarise(chi_not_in_pds =sum(In_PDS0), 
             chi_in_pds = sum(In_PDS1)) %>%
   mutate(perc_chi_in_pds = round(chi_in_pds/(chi_in_pds+chi_not_in_pds)*100,1) )
 
 
 knitr::kable(annualsmr_chi, caption = "Number and proportion of unique CHI in SMR01,01-1E and 04 that are in PDS, by year", 
              col.names = c("Year of admission", "N CHIs not in PDS", "N CHI in PDS", "% unique CHI in PDS"))
```

### Social care CHI not in PDS by year

Social care data also shows an increasing proportion of captured in PDS more recently. Note that the earliest quarter we have data for is 2017/18 Q4, so many people flagged in 2018 will have been diagnosed long before this, and recieving social care before this. This explains both the very high numbers first found in 2018 and the much lower proportion of those CHI found in PDS.

```{r chi_compare_yr_SC,  echo=FALSE}
SC_list <- social_care  %>% select(upi_number, diagnosis_date) %>% 
  left_join(PDS_chi_list, by=c("upi_number" = "pds_chi_number")) %>%
  mutate(inpds = case_when(is.na(inpds)~0, T~inpds)) %>%
  mutate(year_diag = year(diagnosis_date))

annual_SC_chi <-SC_list %>%  group_by(inpds, year_diag) %>% count() %>% 
  pivot_wider(names_from = inpds, values_from = n, names_prefix = "In_PDS")
annual_SC_chi <- annual_SC_chi %>%
   mutate(perc_chi_in_pds = round(In_PDS1/(In_PDS0+In_PDS1)*100,1) )
 

 knitr::kable(annual_SC_chi, caption = "Number and proportion of unique CHI in Social care data that are in PDS, by year of first record in social care", 
              col.names = c("Year dementia first recorded", "N CHIs not in PDS", "N CHI in PDS", "% unique CHI in PDS"))
```
### Care home census CHI not in PDS by year

Social care data also shows an increasing proportion of captured in PDS more recently. Note that the earliest quarter we have data for is 2017/18 Q4, so many people flagged in 2018 will have been diagnosed long before this, and recieving social care before this. This explains both the very high numbers first found in 2018 and the much lower proportion of those CHI found in PDS.

```{r chi_compare_yr_CHC,  echo=FALSE}
##chc
chc_list <- CHC  %>% mutate(year_admit = year(CHC_first_admission)) %>%
  filter(year_admit >=1990) %>%
  select(CHC_UPI_NUMBER, CHC_first_admission) %>% 
  left_join(PDS_chi_list, by=c("CHC_UPI_NUMBER" = "pds_chi_number")) %>%
  mutate(inpds = case_when(is.na(inpds)~0, T~inpds)) 

chc_annual <- chc_list %>% 
  group_by(CHC_UPI_NUMBER) %>% slice(1) %>% ungroup() %>%
  mutate(year_diag = year(CHC_first_admission)) %>%
  mutate(year_diag = case_when(year_diag <= 2014 ~"before 2014", T~as.character(year_diag)))%>%
    ungroup() %>% 
  group_by(inpds, year_diag) %>% count() %>% 
  pivot_wider(names_from = inpds, values_from = n, names_prefix = "In_PDS") 
chc_annual <- chc_annual%>%
   mutate(perc_chi_in_pds = round(In_PDS1/(In_PDS0+In_PDS1)*100,1) )



 knitr::kable(chc_annual, caption = "Number and proportion of unique CHI in Care Home Census data that are in PDS, by year first admission to carehome", 
              col.names = c("Year admission of stay recording dementia", "N CHIs not in PDS", "N CHI in PDS", "% unique CHI in PDS"))
```

### PIS CHI by year

THe PIS data appears to have good capture from 2017, and little before then. FOr this rason, most people in the dataset have thier "first" prescription recorded in 2017. From 2018 onward over 70% of the chi in PIS also appear in PDS. 


```{r chi_compare_yr_pis,  echo=FALSE}
##chc
pis_list <- pis  %>% mutate(year_presc = year(pis_presc_date)) %>%
  filter(year_presc >=1990) %>%
  select(pis_pat_upi_c, year_presc ) %>% 
  left_join(PDS_chi_list, by=c("pis_pat_upi_c" = "pds_chi_number")) %>%
  mutate(inpds = case_when(is.na(inpds)~0, T~inpds)) 

pis_annual <- pis_list %>% 
  group_by(pis_pat_upi_c) %>% slice(1) %>% ungroup() %>%
  mutate(year_diag = year_presc) %>%
#  mutate(year_diag = case_when(year_diag <= 2014 ~"before 2014", T~as.character(year_diag)))%>%
    ungroup() %>% 
  group_by(inpds, year_diag) %>% count() %>% 
  filter(year_diag >=2016) %>%
  pivot_wider(names_from = inpds, values_from = n, names_prefix = "In_PDS") 
pis_annual <- pis_annual%>%
   mutate(perc_chi_in_pds = round(In_PDS1/(In_PDS0+In_PDS1)*100,1) )



 knitr::kable(pis_annual, caption = "Number and proportion of unique CHI in PIS data that are in PDS, by year of first prescription from BNF 4.11", 
              col.names = c("Year of first prescribing", "N CHIs not in PDS", "N CHI in PDS", "% unique CHI in PDS"))
```


### NRS deaths CHI not in PDS by year



```{r chi_compare_yr_death,  echo=FALSE}
death_chi_list <- dementia_deaths %>% 
  mutate(year_diag = year(date_of_death)) %>%
           group_by(upi_number, year_diag) %>% slice(1) %>% ungroup() %>%
  mutate(valid_chi = chi_check(upi_number))

PDS_chi_list <- PDS %>% group_by(pds_chi_number) %>% slice(1) %>%
  mutate(inpds = 1)
  
death_chi_list <-death_chi_list%>% left_join(PDS_chi_list, by=c("upi_number" = "pds_chi_number")) %>%
  mutate(inpds = case_when(is.na(inpds)~0, T~inpds)) 

overall_death_chi <- death_chi_list  %>% group_by(upi_number) %>% slice(1) %>% 
  ungroup() %>% group_by(inpds) %>%count()

annual_death_chi <- death_chi_list %>%  group_by(inpds, year_diag) %>% count() %>% 
  pivot_wider(names_from = inpds, values_from = n, names_prefix = "In_PDS")
 annualsmr_chi <-annual_death_chi %>%
   mutate(year_diag = case_when(year_diag < 2014 ~"before 2014", T~ as.character(year_diag))) %>% 
   group_by(year_diag) %>% 
   summarise(chi_not_in_pds =sum(In_PDS0), 
             chi_in_pds = sum(In_PDS1)) %>%
   mutate(perc_chi_in_pds = round(chi_in_pds/(chi_in_pds+chi_not_in_pds)*100,1) )
 
 
 knitr::kable(annualsmr_chi, caption = "Number and proportion of unique CHI in NRS deaths that are in PDS, by year", 
              col.names = c("Year of admission", "N CHIs not in PDS", "N CHI in PDS", "% unique CHI in PDS"))
```


```{r reverse_chi_compare,  echo=FALSE}
## PDS CHI only in PDS
#PDS_only <- PDS_chi_list %>% 
#  filter(!pds_chi_number %in% smr_chi_list$smr_upi_number) %>%
#  filter(!pds_chi_number %in% CHC$CHC_UPI_NUMBER) %>%
#  filter(!pds_chi_number %in% pis_chi$pis_pat_upi_c) %>%
#  filter(!pds_chi_number %in% SC_first$SC_chi) 


```

## Summary of dementia types

PDS provides the most detail of dementia subtype, classifying dementia into 7 categories (including a category for those where the subtype has not been determined yet). The prescribing, care home and social care datasets only flag dementia, with no further detail. However the SMR sources and the NRS deaths use ICD10 codes and so can potentially provide more details on the subtype of dementia present. 



```{r dem_types,  echo=FALSE}
#PDS <- readRDS("/PHI_conf/Dementia_Index/data/extracts/PDS_clean.rds") 
pds_types <- PDS %>% group_by(pds_dementia_subtype) %>%count() %>%
  rename(PDS = n,dementia_subtype= pds_dementia_subtype)
smr_types <- smr_all %>% group_by(smr_dementia_subtype_1) %>%count() %>%
  mutate(smr_dementia_subtype_1 = 
           case_when(smr_dementia_subtype_1=="07 yet to be determined"~ "07 Yet to be determined", 
                     T~smr_dementia_subtype_1)) %>%
  rename(first_SMRA_type = n,dementia_subtype = smr_dementia_subtype_1 ) 
smr_types2 <- smr_all %>% group_by(smr_dementia_subtype_2) %>%count() %>%
  filter(!is.na(smr_dementia_subtype_2)) %>%
  rename(second_SMRA_type = n,dementia_subtype = smr_dementia_subtype_2 ) 


death_types <- dementia_deaths %>%
  group_by(NRSdeaths_dementia_subtype_1) %>% count() %>%
   rename(NRSdeaths = n, dementia_subtype = NRSdeaths_dementia_subtype_1) %>%
  mutate(dementia_subtype = 
           case_when(dementia_subtype=="07 yet to be determined"~ "07 Yet to be determined", 
                     T~dementia_subtype))

df <- full_join( pds_types,smr_types)
df <- full_join( df,smr_types2)
df <- full_join(df, death_types)

totals <- c("Total", colSums(df[2:5], na.rm=T))
names(totals) <- names(df)
totals <-(as.data.frame(t(totals)))
totals <- totals %>% mutate(PDS =as.numeric(PDS), 
                            first_SMRA_type = as.numeric(first_SMRA_type), 
                            second_SMRA_type = as.numeric(second_SMRA_type), 
                            NRSdeaths = as.numeric(NRSdeaths))

df <- rbind(df , totals)
knitr::kable(df %>% arrange(dementia_subtype), caption = "Dementia subtype frequency in PDS, SMR records and NRS deaths")
```



### Care home census
```{r chc_flags,  echo=FALSE}
CHC_MD <- CHC %>% filter(CHC_dementia_subtype=="07 Yet to be determined")
CHC_NMD <- CHC %>% filter(CHC_dementia_subtype=="99 Suspected dementia")

both <- CHC %>% filter(CHC_UPI_NUMBER %in% CHC_NMD$CHC_UPI_NUMBER & CHC_UPI_NUMBER %in% CHC_MD$CHC_UPI_NUMBER)
n_nmd <-nrow(CHC_NMD)
n_md <- length(unique(CHC_MD$CHC_UPI_NUMBER))

n_nmd_only <- n_nmd-nrow(both)

chc_unique_chi <- length(unique(CHC$CHC_UPI_NUMBER))


perc_nmd_only <- round((n_nmd_only/chc_unique_chi)*100,1)
perc_medic <- round(n_md/chc_unique_chi*100,1)
```
*add table of number with MD/NMD dementia*


`r n_md` (`r perc_medic`%) of persons were marked as having medically diagnosed dementia (they may have also been marked as NMD before or after the medical diagnosis)

`r n_nmd_only` (`r perc_nmd_only`%) people were *only* marked as NMD dementia


