---
title: "Dementia Analysis Dashboard"

output:
  flexdashboard::flex_dashboard:
    orientation: rows
    logo: phs monochrome logo.png
    vertical_layout: scroll
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Notes {data-icon="fa-info-circle"}

#### <span style="color:red;text-decoration:underline;font-size:28px">**Not for onward distribution.**</span>

#### <span style="text-decoration:underline">**Dementia Index Recorded Prevalence**</span>
#### **Scotland and HSCP Level 2017/18 - 2023/24** 

Dashboard showing recorded prevalence and death figures for people aged 18+ at Scotland and HSCP level with age, sex, SIMD and urban rural classification breakdowns.

Please note that prevalence figures do not capture 100% of diagnoses in Scotland as these figures do not include data of recorded diagnosis from GP systems. Therefore figures are likely lower than the true number and should only be used as an estimate.

#### <span style="text-decoration:underline">**Pages**</span>

 - [Recorded Dementia Prevalence](#recorded-dementia-prevalence) - Recorded prevalence broken down by age, sex and data sources where dementia diagnosis appeared
 - [Deaths](#deaths) - Deaths broken down by age group and sex (65+ only)
 - [GP Dementia Prevalence Comparison](#gp-dashboard-comparison) - Comparison with GP disease prevalence figures for dementia, broken down by age and sex
 - [SIMD](#simd) - Recorded prevalence breakdowns by SIMD2020v2 quintiles
 - [Urban Rural Classification](#urban-rural-classification) - Recorded Prevalence broken down by UR8 & UR6 classification (2022)

#### <span style="text-decoration:underline">**Data Sources**</span>

 - SMR01, SMR01-E, SMR04
 - Dementia Post Diagnostic Support (PDS)
 - Care Home Census
 - PIS
 - Social Care

#### <span style="text-decoration:underline">**Dementia Index Contacts**</span>

If you have any questions about this output, or would like to enquire about using the data you can contact any of the following:

 - Jade Carruthers | Principal Information Analyst - <a href="mailto:jade.carruthers@phs.scot">Jade.carruthers@phs.scot</a>
 - Kwaku Osei-Oppong | Senior Information Analyst - <a href="mailto:kwaku.osei-oppong@phs.scot">kwaku.osei-oppong@phs.scot</a>
 - Jack Donaghy | Senior Information Analyst - <a href="mailto:jack.donaghy2@phs.scot">jack.donaghy2@phs.scot</a>


# Recorded Dementia Prevalence {data-icon="fa-chart-line"}

Inputs {.sidebar}
-----------------------------------------------------------------------

```{r filter_section}
# prev_filter_select
filter_select("area_select1", "Select an Area",
                                    shared_prev_65plus_chart, group = ~area, allLevels = T, multiple = F)
```

**Notes:**

<p>There is a reduction in numbers due to the COVID period where fewer hospital admissions were recorded. There was no care home census carried out in 2019/20 and from the period of 2018/19 to 2021/22 there was a decrease of 70% in admissions for respite residents.</p>

<p>**Dementia Service Contacts & Patient Total** - this table shows a total count of all patients by each source they are found in, so a patient may appear in more than one source. The patient total count is a count of each person who appeared in any source</p>

Row
-----------------------------------------------------------------------

### Dementia Count - Aged 65+
```{r 65+ table}
datatable(shared_prev_65plus_table,
          options = list(dom = 't'),
          rownames = F,
          selection = 'none',
          colnames = c("Year", "Area", "Female", "Male", "Total")) %>%
  formatCurrency(c('female', 'male', 'total'),currency = "", interval = 3, mark = ",", digits = 0)
```


### Dementia Count in Area - Aged 65+
```{r 65+ chart}
plot_ly(shared_prev_65plus_chart, x = ~year, y = ~individuals, meta = ~gender, hovertext = ~format(individuals, big.mark=',', scientific = FALSE), type = 'bar', 
          color = ~gender, colors = c("#9B4393", "#0078D4"),
          hovertemplate = paste0("%{x} <br>", "%{meta} <br>", "Individuals: %{hovertext} <br>")) %>% 
    layout(clickmode = "none", showlegend = TRUE, margin = list(l = -5, b = 10, t = 40),
           xaxis = list(title = "Financial Year", showline = FALSE, linecolor = "#b3b3b3",
                        categoryorder = "trace", tickangle = -45, titlefont = list(size = 16)),
           yaxis = list(title = "Count", linecolor = "#b3b3b3", titlefont = list(size = 16))
    ) %>%
    config(displayModeBar = TRUE, doubleClick = F,
           modeBarButtonsToRemove = list('select2d', 'lasso2d', 'zoomIn2d', 
                                         'zoomOut2d', 'autoScale2d', 
                                         'toggleSpikelines', 
                                         'hoverCompareCartesian', 
                                         'hoverClosestCartesian', 'toImage'), 
           displaylogo = F, editable = F)
```

Row {.tabset}
-----------------------------------------------------------------------

### Dementia Rate per 100,000 - Aged 65+ Chart 
```{r rate chart}
plot_ly(shared_prev_65plus_chart, x = ~year, y = ~rate_100000, meta = ~gender, hovertext = ~format(round_half_up(rate_100000, digits = 1), big.mark=',', scientific = FALSE), 
          type = 'scatter', mode = 'lines', color = ~gender, colors = c("#9B4393", "#0078D4"),
          hovertemplate = paste0("%{x} <br>", "%{meta} <br>", "Rate: %{hovertext} <br>")) %>% 
    layout(clickmode = "none", showlegend = TRUE, margin = list(l = -5, b = 10, t = 40),
           xaxis = list(title = "Financial Year", showline = FALSE, linecolor = "#b3b3b3",
                        range = list(-0.4, length(sort(unique(prevalence_all_65plus$year)))-1),
                        categoryorder = "trace", titlefont = list(size = 16)),
           yaxis = list(title = "Prevalence rate per 100,000 pop (Aged 65+)", titlefont = list(size = 16), rangemode = "tozero")
    ) %>%
    config(displayModeBar = TRUE, doubleClick = F,
           modeBarButtonsToRemove = list('select2d', 'lasso2d', 'zoomIn2d', 
                                         'zoomOut2d', 'autoScale2d', 
                                         'toggleSpikelines', 
                                         'hoverCompareCartesian', 
                                         'hoverClosestCartesian', 'toImage'), 
           displaylogo = F, editable = F)
```

### Dementia Count and Rate per 100,000 - Aged 65+ Table
```{r rate table}
# datatable(shared_prev_65plus_rate_table, options = list(dom = 't'),
#           rownames = F,
#           selection = 'none',
#           colnames = c("Area", "Gender", fy_dates)) %>%
#   formatCurrency(fy_dates, currency = "", interval = 3, mark = ",", digits = 0)

count_rate_cols <- colnames(prevalence_all_count_rates)[4:7]

datatable(shared_prev_65plus_count_rate_table, options = list(dom = 't'),
          rownames = F,
          colnames = c("Year", "Area", count_rate_cols),
          selection = 'none') %>%
  formatCurrency(count_rate_cols, currency = "", interval = 3, mark = ",", digits = 0)
```


Row {.tabset}
-----------------------------------------------------------------------

### Dementia Age Group Split - Count
```{r age totals}
datatable(shared_prev_all_total, options = list(dom = 't'),
          rownames = F,
          selection = 'none',
          colnames = c("Year", "Area", age_order, "Total")) %>%
  formatCurrency(c(age_order, 'Total'), currency = "", interval = 3, mark = ",", digits = 0)
```

### Dementia Age Group Split - per 100,000 population
```{r age totals rate}
datatable(shared_prev_all_rate, options = list(dom = 't'),
          rownames = F,
          selection = 'none',
          colnames = c("Year", "Area", age_order)) %>%
  formatCurrency(age_order, currency = "", interval = 3, mark = ",", digits = 0)
```

Row
-----------------------------------------------------------------------

### Dementia Service Contacts & Patient Total 
```{r totals by source}
datatable(shared_prev_source, options = list(dom = 't'),
          rownames = F,
          selection = 'none',
          colnames = c("Year", "Area", "Care Home Census", 
                       colnames(prev_all_source)[4:8], 
                       "Social Care", "Patient Total")) %>%
  formatCurrency(colnames(prev_all_source)[3:10], currency = "", interval = 3, mark = ",", digits = 0)
```

# Deaths {data-icon="fa-chart-line"}

Inputs {.sidebar}
-----------------------------------------------------------------------

```{r filter_section deaths with}
# deaths_filter_select
filter_select("area_select_deaths", "Select an Area",
                                    shared_all_deaths_with_table, group = ~area, allLevels = T, multiple = F)
```

**Notes:**

**Deaths of those with a dementia diagnosis** - These figures show deaths for those who had a diagnosis of dementia either prior to death or as a cause of death and those who had no dementia diagnosis within the source datasets. The percent figures show the percent of all deaths for the year that were attributed to those aged 65+ who also had a diagnosis of dementia.

**NRS Dementia Deaths** - These figures show all deaths caused by dementia and deaths mentioning dementia on the death certificate. Deaths caused by dementia are where a dementia code (F01, F03 or G30) is determined as the underlying cause of death. Deaths mentioning dementia are any death where dementia is either an underlying or contributing cause of death. More information on NRS definitions and methodology can be found in their <a href="https://www.nrscotland.gov.uk/publications/alzheimer-s-and-other-dementia-deaths/#" target="_blank">Alzheimer’s and other dementia deaths publication</a>

Row {.tabset}
-----------------------------------------------------------------------

### Deaths of those with a dementia diagnosis (65+)
```{r all deaths with 65+}
datatable(shared_all_deaths_with_table, options = list(dom = 't'),
          rownames = F,
          selection = 'none',
          colnames = c("Year", "Area", "Dementia Cohort Deaths", "Deaths with no dementia diagnosis", "Percent Dementia")) %>%
  formatCurrency(c('dementia', 'non_dementia'), currency = "", interval = 3, mark = ",", digits = 0)# %>% formatPercentage('proportion_to_dementia')
```

### NRS Dementia Deaths (65+)
```{r nrs deaths 65+}
datatable(shared_nrs_deaths, options = list(dom = 't'),
          rownames = F,
          selection = 'none',
          colnames = c("Year", "Area", colnames(nrs_deaths)[3:5])) %>% 
  formatCurrency(c(colnames(nrs_deaths))[3:4], currency = "", interval = 3, mark =  ", ", digits = 0)
```

Row {.tabset}
-----------------------------------------------------------------------

### Percent of all deaths attributed to dementia cohort - age group (65 & over)
```{r proportion with chart age}
plot_ly(shared_deaths_with_chart_age, x = ~year, y = ~proportion, split = ~age_group, meta = ~age_group, hovertext = ~proportion, type = 'bar', 
        color = ~age_group, colors = unname(phsstyles::phs_colour_values)[1:6],
        hovertemplate = paste0("%{x} <br>", "Age Group: %{meta} <br>", "Percent of deaths attributed to dementia cohort: %{hovertext}% <br>")) %>% 
  layout(clickmode = "none", showlegend = TRUE, margin = list(l = -5, b = 10, t = 40),
         xaxis = list(title = "Financial Year", showline = FALSE, linecolor = "#b3b3b3",
                      categoryorder = "trace", tickangle = -45, titlefont = list(size = 16)),
         yaxis = list(title = "%", linecolor = "#b3b3b3", titlefont = list(size = 16))
  ) %>%
  config(displayModeBar = TRUE, doubleClick = F,
         modeBarButtonsToRemove = list('select2d', 'lasso2d', 'zoomIn2d', 
                                       'zoomOut2d', 'autoScale2d', 
                                       'toggleSpikelines', 
                                       'hoverCompareCartesian', 
                                       'hoverClosestCartesian', 'toImage'), 
         displaylogo = F, editable = F)
```

### Table
```{r proportion with table age}
datatable(shared_deaths_with_chart_age, options = list(dom = 't',
                         pageLength = 42),
          rownames = F,
          selection = 'none',
          colnames = c("Year", "Area", "Age Group", "Dementia Cohort Deaths", "Deaths with no dementia diagnosis", "Percent Dementia")) %>%
  formatCurrency(c('dementia', 'non_dementia'), currency = "", interval = 3, mark = ",", digits = 0)
```

Row {.tabset}
-----------------------------------------------------------------------

### Percent of all deaths attributed to dementia cohort - gender (65+ only)
```{r proportion with chart gender}
plot_ly(shared_deaths_with_chart_gender, x = ~year, y = ~proportion, split = ~gender, meta = ~gender, hovertext = ~proportion, type = 'bar', 
        color = ~gender, colors = c("#9B4393", "#0078D4"),
        hovertemplate = paste0("%{x} <br>", "Gender: %{meta} <br>", "Percent of deaths attributed to dementia cohort: %{hovertext}% <br>")) %>% 
  layout(clickmode = "none", showlegend = TRUE, margin = list(l = -5, b = 10, t = 40),
         xaxis = list(title = "Financial Year", showline = FALSE, linecolor = "#b3b3b3",
                      categoryorder = "trace", tickangle = -45, titlefont = list(size = 16)),
         yaxis = list(title = "%", linecolor = "#b3b3b3", titlefont = list(size = 16))
  ) %>%
  config(displayModeBar = TRUE, doubleClick = F,
         modeBarButtonsToRemove = list('select2d', 'lasso2d', 'zoomIn2d', 
                                       'zoomOut2d', 'autoScale2d', 
                                       'toggleSpikelines', 
                                       'hoverCompareCartesian', 
                                       'hoverClosestCartesian', 'toImage'), 
         displaylogo = F, editable = F)
```

### Table
```{r proportion with table gender}
datatable(shared_deaths_with_chart_gender, options = list(dom = 't', 
                         pageLength = 42),
          rownames = F,
          selection = 'none',
          colnames = c("Year", "Area", "Gender", "Dementia Cohort Deaths", "Deaths with no dementia diagnosis", "Percent Dementia")) %>%
  formatCurrency(c('dementia', 'non_dementia'), currency = "", interval = 3, mark = ",", digits = 0)
```

# GP Dashboard Comparison {data-icon="fa-chart-line"}

Inputs {.sidebar}
-----------------------------------------------------------------------

```{r filter_section GP comparison}
# gp_filter_select
filter_select("area_select2", "Select an Area",
                                    shared_prev_comparison_65plus_table, group = ~area, allLevels = T, multiple = F)
```

**Notes:**

The figures included in this tab show GP recorded prevalence from the dementia index compared to those from the <a href="https://www.publichealthscotland.scot/publications/general-practice-disease-prevalence-data-visualisation/general-practice-disease-prevalence-visualisation-8-july-2025/" target="_blank">GP Disease Prevalence Dashboard</a> which contains data extracted from General Practices in Scotland through the Primary Care Information Service (PCIS).

Rates from GP data are reported per 100, therefore rates for the index have been adjusted accordingly on this page.

Row
-----------------------------------------------------------------------

### Dementia Count Aged 65+ Comparison
```{r 65 plus comparison}
datatable(shared_prev_comparison_65plus_table, options = list(dom = 't'),
          rownames = F,
          selection = 'none',
          colnames = c("Year", "Area", "Source", "Female", "Male", "Total")) %>%
  formatCurrency(c('Female', 'Male', 'Total'), currency = "", interval = 3, mark = ",", digits = 0)
```

### Dementia Count in Area 65+ Comparison
```{r 65 plus chart comparison}
plot_ly(shared_prev_comparison_65plus_chart, 
        x = ~year, y = ~individuals, meta = ~group, hovertext = ~format(individuals, big.mark=',', scientific = FALSE), type = 'bar', 
        color = ~group, split = ~source, colors = unname(phsstyles::phs_colour_values)[1:4],
        hovertemplate = paste0("%{x} <br>", "%{meta} <br>", "Individuals: %{hovertext} <br>")) %>% 
  layout(clickmode = "none", showlegend = TRUE, margin = list(l = -5, b = 10, t = 40),
         xaxis = list(title = "Financial Year", showline = FALSE, linecolor = "#b3b3b3",
                      categoryorder = "trace", tickangle = -45, titlefont = list(size = 16)),
         yaxis = list(title = "Count", linecolor = "#b3b3b3", titlefont = list(size = 16))
  ) %>%
  config(displayModeBar = TRUE, doubleClick = F,
         modeBarButtonsToRemove = list('select2d', 'lasso2d', 'zoomIn2d', 
                                       'zoomOut2d', 'autoScale2d', 
                                       'toggleSpikelines', 
                                       'hoverCompareCartesian', 
                                       'hoverClosestCartesian', 'toImage'), 
         displaylogo = F, editable = F)
```

Row {.tabset}
-----------------------------------------------------------------------

### Age Group Split Comparisons - Counts
```{r age group comparison}
datatable(shared_prev_comparsion_ages_source_numbers, options = list(dom = 't'),
          rownames = F,
          selection = 'none',
          colnames = c("Year", "Area", "Source", age_order2, "Total")) %>%
  formatCurrency(c(age_order2, "Total"), currency = "", interval = 3, mark = ",", digits = 0)
```

### Age Group Split Comparisons - Rate per 100 
```{r age groups over 65 comparison}
datatable(shared_prev_comparsion_ages_source_rates, options = list(dom = 't'),
          rownames = F,
          selection = 'none',
          colnames = c("Year", "Area", "Source", age_order_gp)) %>%
  formatCurrency(c(age_order_gp), currency = "", interval = 3, mark = ",", digits = 1)
```

# SIMD {data-icon="fa-chart-line"}

Inputs {.sidebar}
-----------------------------------------------------------------------

**SIMD 65+ Age Group**

```{r filter_section SIMD}
filter_select("area_select_simd", "Select an Area",
                                    shared_simd_rates_65plus, group = ~area, allLevels = T, multiple = F)
```

***

<ul>
  <li>[**65+ Age Group**](#simd)</li>
  <li>[All Age Groups](#simd-all)</li>
</ul>

***

Row {.tabset}
-----------------------------------------------------------------------

### 65+ SIMD Chart
```{r 65+ simd numbers chart}
plot_ly(shared_simd_total_65plus_ch, 
        x = ~year, y = ~individuals, 
        meta = ~simd2020v2_sc_quintile, split = ~simd2020v2_sc_quintile,
        hovertext = ~format(individuals, big.mark=',', scientific = FALSE), type = 'bar', 
        color = ~simd2020v2_sc_quintile, colors = unname(phsstyles::phs_colour_values)[1:5],
        hovertemplate = paste0("%{x} <br>", "%{meta} <br>", "Individuals: %{hovertext} <br>")) %>% 
  layout(clickmode = "none", showlegend = TRUE, margin = list(l = -5, b = 10, t = 40),
         xaxis = list(title = "Financial Year", showline = FALSE, linecolor = "#b3b3b3",
                      categoryorder = "trace", tickangle = -45, titlefont = list(size = 16)),
         yaxis = list(title = "Count", linecolor = "#b3b3b3", titlefont = list(size = 16))
  ) %>%
  config(displayModeBar = TRUE, doubleClick = F,
         modeBarButtonsToRemove = list('select2d', 'lasso2d', 'zoomIn2d', 
                                       'zoomOut2d', 'autoScale2d', 
                                       'toggleSpikelines', 
                                       'hoverCompareCartesian', 
                                       'hoverClosestCartesian', 'toImage'), 
         displaylogo = F, editable = F)
```

### 65+ SIMD table
```{r 65+ simd numbers table}
datatable(shared_simd_total_65plus,# extensions = 'Scroller', 
          options = list(dom = 't',
                         pageLength = 42,
                         columnDefs = list(list(visible=FALSE, targets=c(7)))),
          rownames = F,
          fillContainer = T,
          selection = 'none',
          colnames = c("Year", "Area", simd_65_cols, "key")) %>%
    formatCurrency(simd_65_cols, currency = "", interval = 3, mark = ",", digits = 0)
```

Row {.tabset}
-----------------------------------------------------------------------

### Dementia - per 100,000 population 65+ chart
```{r all age simd totals}
plot_ly(shared_simd_rates_65plus_ch, 
        x = ~year, y = ~round_half_up(rate, digits = 1), 
        meta = ~simd2020v2_sc_quintile, split = ~simd2020v2_sc_quintile,
        hovertext = ~format(round_half_up(rate, digits = 1), big.mark=',', scientific = FALSE), type = 'bar', 
        color = ~simd2020v2_sc_quintile, colors = unname(phsstyles::phs_colour_values)[1:5],
        hovertemplate = paste0("%{x} <br>", "%{meta} <br>", "Rate: %{hovertext} <br>")) %>% 
  layout(clickmode = "none", showlegend = TRUE, margin = list(l = -5, b = 10, t = 40),
         xaxis = list(title = "Financial Year", showline = FALSE, linecolor = "#b3b3b3",
                      categoryorder = "trace", tickangle = -45, titlefont = list(size = 16)),
         yaxis = list(title = "Rate per 100,000", linecolor = "#b3b3b3", titlefont = list(size = 16))
  ) %>%
  config(displayModeBar = TRUE, doubleClick = F,
         modeBarButtonsToRemove = list('select2d', 'lasso2d', 'zoomIn2d', 
                                       'zoomOut2d', 'autoScale2d', 
                                       'toggleSpikelines', 
                                       'hoverCompareCartesian', 
                                       'hoverClosestCartesian', 'toImage'), 
         displaylogo = F, editable = F)
```


### Dementia - per 100,000 population 65+ table
```{r 65+ simd rates chart}
datatable(shared_simd_rates_65plus, options = list(dom = 't',
                                                   columnDefs = list(list(visible=FALSE, targets=c(7)))),
          rownames = F,
          fillContainer = TRUE,
          selection = 'none',
          colnames = c("Year", "Area", simd_65_cols, "key")) %>%
    formatCurrency(simd_65_cols, currency = "", interval = 3, mark = ",", digits = 0)

```

# SIMD All {.hidden}

Inputs {.sidebar}
-----------------------------------------------------------------------

**SIMD All Age Groups**

```{r filter_section SIMD hidden}
filter_select("area_select_simd2", "Select an Area",
                                    shared_simd_rates_ch, group = ~area, allLevels = T, multiple = F)

filter_select("year_select_simd2", "Select a Financial Year",
                                    shared_simd_rates_ch, group = ~year, allLevels = T, multiple = F)
```

***

<ul>
  <li>[65+ Age Group](#simd)</li>
  <li>[**All Age Groups**](#simd-all)</li>
</ul>

***

Row {.tabset}
-----------------------------------------------------------------------

### SIMD Chart
```{r simd numbers chart hidden}
plot_ly(shared_simd_total_ch, 
        x = ~age_group, y = ~individuals, 
        meta = ~simd2020v2_sc_quintile, split = ~simd2020v2_sc_quintile,
        hovertext = ~format(individuals, big.mark=',', scientific = FALSE), type = 'bar', 
        color = ~simd2020v2_sc_quintile, colors = unname(phsstyles::phs_colour_values)[1:5],
        hovertemplate = paste0("%{x} <br>", "%{meta} <br>", "Individuals: %{hovertext} <br>")) %>% 
  layout(clickmode = "none", showlegend = TRUE, margin = list(l = -5, b = 10, t = 40),
         xaxis = list(title = "Age Group", showline = FALSE, linecolor = "#b3b3b3",
                      categoryorder = "trace", tickangle = -45, titlefont = list(size = 16)),
         yaxis = list(title = "Count", linecolor = "#b3b3b3", titlefont = list(size = 16))
  ) %>%
  config(displayModeBar = TRUE, doubleClick = F,
         modeBarButtonsToRemove = list('select2d', 'lasso2d', 'zoomIn2d', 
                                       'zoomOut2d', 'autoScale2d', 
                                       'toggleSpikelines', 
                                       'hoverCompareCartesian', 
                                       'hoverClosestCartesian', 'toImage'), 
         displaylogo = F, editable = F)
```

### SIMD table
```{r simd numbers table hidden}
datatable(shared_simd_total,# extensions = 'Scroller', 
          options = list(dom = 't',
                         pageLength = 42,
                         columnDefs = list(list(visible=FALSE, targets=c(11)))),
          rownames = F,
          fillContainer = T,
          selection = 'none',
          colnames = c("Year", "Area", "SIMD Quintile", age_order, "key")) %>%
    formatCurrency(age_order, currency = "", interval = 3, mark = ",", digits = 0)
```

Row {.tabset}
-----------------------------------------------------------------------

### Dementia Age Group SIMD Split - per 100,000 population chart
```{r all age simd totals hidden}
plot_ly(shared_simd_rates_ch,
        x = ~age_group, y = ~round_half_up(rate, digits = 1),
        meta = ~simd2020v2_sc_quintile, split = ~simd2020v2_sc_quintile,
        hovertext = ~format(round_half_up(rate, digits = 1), big.mark=',', scientific = FALSE), type = 'bar',
        color = ~simd2020v2_sc_quintile, colors = unname(phsstyles::phs_colour_values)[1:5],
        hovertemplate = paste0("%{x} <br>", "%{meta} <br>", "Rate: %{hovertext} <br>")) %>%
  layout(clickmode = "none", showlegend = TRUE, margin = list(l = -5, b = 10, t = 40),
         xaxis = list(title = "Age Group", showline = FALSE, linecolor = "#b3b3b3",
                      categoryorder = "trace", tickangle = -45, titlefont = list(size = 16)),
         yaxis = list(title = "Rate per 100,000", linecolor = "#b3b3b3", titlefont = list(size = 16))
  ) %>%
  config(displayModeBar = TRUE, doubleClick = F,
         modeBarButtonsToRemove = list('select2d', 'lasso2d', 'zoomIn2d',
                                       'zoomOut2d', 'autoScale2d',
                                       'toggleSpikelines',
                                       'hoverCompareCartesian',
                                       'hoverClosestCartesian', 'toImage'),
         displaylogo = F, editable = F)
```


### Dementia Age Group SIMD Split - per 100,000 population table
```{r simd rates table hidden}
datatable(shared_simd_rates, options = list(dom = 't',
                                            columnDefs = list(list(visible=FALSE, targets=c(11)))),
          rownames = F,
          fillContainer = TRUE,
          selection = 'none',
          colnames = c("Year", "Area", "SIMD Quintile", age_order, "key")) %>%
    formatCurrency(age_order, currency = "", interval = 3, mark = ",", digits = 0)

```

# Urban Rural Classification {data-icon="fa-chart-line"}

Inputs {.sidebar}
-----------------------------------------------------------------------

**URC 65+ Age Group**

```{r filter_section urc}
filter_select("area_select_urc", "Select an Area",
                                    shared_urc8_65plus_ch, group = ~area, allLevels = T, multiple = F)
```

***

<ul>
  <li>[**65+ Age Group**](#urban-rural-classification)</li>
  <li>[All Age Groups](#urban-rural-classification-all)</li>
</ul>

***

Row {.tabset}
-----------------------------------------------------------------------

### 65+ UR8 Chart
```{r 65+ urc8 numbers chart}
plot_ly(shared_urc8_65plus_ch, 
        x = ~year, y = ~individuals, 
        meta = ~ur8_name, split = ~ur8_name,
        hovertext = ~format(individuals, big.mark=',', scientific = FALSE), type = 'bar', 
        color = ~ur8_name, colors = unname(phsstyles::phs_colour_values)[1:8],
        hovertemplate = paste0("%{x} <br>", "%{meta} <br>", "Individuals: %{hovertext} <br>")) %>% 
  layout(clickmode = "none", showlegend = TRUE, margin = list(l = -5, b = 10, t = 40),
         xaxis = list(title = "Financial Year", showline = FALSE, linecolor = "#b3b3b3",
                      categoryorder = "trace", tickangle = -45, titlefont = list(size = 16)),
         yaxis = list(title = "Count", linecolor = "#b3b3b3", titlefont = list(size = 16))
  ) %>%
  config(displayModeBar = TRUE, doubleClick = F,
         modeBarButtonsToRemove = list('select2d', 'lasso2d', 'zoomIn2d', 
                                       'zoomOut2d', 'autoScale2d', 
                                       'toggleSpikelines', 
                                       'hoverCompareCartesian', 
                                       'hoverClosestCartesian', 'toImage'), 
         displaylogo = F, editable = F)
```

### 65+ UR8 table
```{r 65+ urc8 numbers table}
datatable(shared_urc8_65plus_table, 
          options = list(dom = 't',
                         pageLength = 42,
                         columnDefs = list(list(visible=FALSE, targets=c(10)))),
          rownames = F,
          fillContainer = T,
          selection = 'none',
          colnames = c("Year", "Area", urc8_cols, "key")) %>%
    formatCurrency(urc8_cols, currency = "", interval = 3, mark = ",", digits = 0)
```


### 65+ UR6 Chart
```{r 65+ urc6 numbers chart}
plot_ly(shared_urc6_65plus_ch, 
        x = ~year, y = ~individuals, 
        meta = ~ur6_name, split = ~ur6_name,
        hovertext = ~format(individuals, big.mark=',', scientific = FALSE), type = 'bar', 
        color = ~ur6_name, colors = unname(phsstyles::phs_colour_values)[1:6],
        hovertemplate = paste0("%{x} <br>", "%{meta} <br>", "Individuals: %{hovertext} <br>")) %>% 
  layout(clickmode = "none", showlegend = TRUE, margin = list(l = -5, b = 10, t = 40),
         xaxis = list(title = "Financial Year", showline = FALSE, linecolor = "#b3b3b3",
                      categoryorder = "trace", tickangle = -45, titlefont = list(size = 16)),
         yaxis = list(title = "Count", linecolor = "#b3b3b3", titlefont = list(size = 16))
  ) %>%
  config(displayModeBar = TRUE, doubleClick = F,
         modeBarButtonsToRemove = list('select2d', 'lasso2d', 'zoomIn2d', 
                                       'zoomOut2d', 'autoScale2d', 
                                       'toggleSpikelines', 
                                       'hoverCompareCartesian', 
                                       'hoverClosestCartesian', 'toImage'), 
         displaylogo = F, editable = F)
```

### 65+ UR6 table
```{r 65+ urc6 numbers table}
datatable(shared_urc6_65plus_table, 
          options = list(dom = 't',
                         pageLength = 42,
                         columnDefs = list(list(visible=FALSE, targets=c(8)))),
            rownames = F,
            fillContainer = T,
            selection = 'none',
          colnames = c("Year", "Area", urc6_cols, "key")) %>%
    formatCurrency(urc6_cols, currency = "", interval = 3, mark = ",", digits = 0)
```

Row {.tabset}
-----------------------------------------------------------------------

### Dementia per 100,000 population UR8 65+ chart
```{r 65+ urc8 rate chart}
plot_ly(shared_urc8_65plus_rate_ch,
        x = ~year, y = ~rate,
        meta = ~ur8_name, split = ~ur8_name,
        hovertext = ~format(round_half_up(rate, digits = 1), big.mark=',', scientific = FALSE), type = 'bar',
        color = ~ur8_name, colors = unname(phsstyles::phs_colour_values)[1:8],
        hovertemplate = paste0("%{x} <br>", "%{meta} <br>", "Rate per 100,000: %{hovertext} <br>")) %>% 
  layout(clickmode = "none", showlegend = TRUE, margin = list(l = -5, b = 10, t = 40),
         xaxis = list(title = "Financial Year", showline = FALSE, linecolor = "#b3b3b3",
                      categoryorder = "trace", tickangle = -45, titlefont = list(size = 16)),
         yaxis = list(title = "%", linecolor = "#b3b3b3", titlefont = list(size = 16))
  ) %>%
  config(displayModeBar = TRUE, doubleClick = F,
         modeBarButtonsToRemove = list('select2d', 'lasso2d', 'zoomIn2d', 
                                       'zoomOut2d', 'autoScale2d', 
                                       'toggleSpikelines', 
                                       'hoverCompareCartesian', 
                                       'hoverClosestCartesian', 'toImage'), 
         displaylogo = F, editable = F)
```

### Dementia per 100,000 population UR8 65+ table
```{r 65+ urc8 rate table}
datatable(shared_urc8_65plus_rate_table, 
          options = list(dom = 't',
                         pageLength = 42,
                         columnDefs = list(list(visible=FALSE, targets=c(10)))),
            rownames = F,
            fillContainer = T,
            selection = 'none',
          colnames = c("Year", "Area", urc8_cols, "key")) %>%
    formatCurrency(urc8_cols, currency = "", interval = 3, mark = ",", digits = 1)
```

### Dementia per 100,000 population UR6 65+ chart
```{r 65+ urc6 rate chart}
plot_ly(shared_urc6_65plus_rate_ch,
        x = ~year, y = ~rate,
        meta = ~ur6_name, split = ~ur6_name,
        hovertext = ~format(round_half_up(rate, digits = 1), big.mark=',', scientific = FALSE), type = 'bar',
        color = ~ur6_name, colors = unname(phsstyles::phs_colour_values)[1:6],
        hovertemplate = paste0("%{x} <br>", "%{meta} <br>", "Rate per 100,000: %{hovertext} <br>")) %>% 
  layout(clickmode = "none", showlegend = TRUE, margin = list(l = -5, b = 10, t = 40),
         xaxis = list(title = "Financial Year", showline = FALSE, linecolor = "#b3b3b3",
                      categoryorder = "trace", tickangle = -45, titlefont = list(size = 16)),
         yaxis = list(title = "%", linecolor = "#b3b3b3", titlefont = list(size = 16))
  ) %>%
  config(displayModeBar = TRUE, doubleClick = F,
         modeBarButtonsToRemove = list('select2d', 'lasso2d', 'zoomIn2d', 
                                       'zoomOut2d', 'autoScale2d', 
                                       'toggleSpikelines', 
                                       'hoverCompareCartesian', 
                                       'hoverClosestCartesian', 'toImage'), 
         displaylogo = F, editable = F)
```

### Dementia per 100,000 population UR6 65+ table
```{r 65+ urc6 rate table}
datatable(shared_urc6_65plus_rate_table, 
          options = list(dom = 't',
                         pageLength = 42,
                         columnDefs = list(list(visible=FALSE, targets=c(8)))),
            rownames = F,
            fillContainer = T,
            selection = 'none',
          colnames = c("Year", "Area", urc6_cols, "key")) %>%
    formatCurrency(urc6_cols, currency = "", interval = 3, mark = ",", digits = 1)
```

# Urban Rural Classification All {.hidden}

Inputs {.sidebar}
-----------------------------------------------------------------------

**URC All Age Groups**

```{r filter_section urc all age}
filter_select("area_select_urc2", "Select an Area",
                                    shared_urc8_all_ch, group = ~area, allLevels = T, multiple = F)
filter_select("year_select_urc", "Select a Financial Year",
                                    shared_urc8_all_ch, group = ~year, allLevels = T, multiple = F)
```

***

<ul>
  <li>[65+ Age Group](#urban-rural-classification)</li>
  <li>[**All Age Groups**](#urban-rural-classification-all)</li>
</ul>

***

Row {.tabset}
-----------------------------------------------------------------------

### UR8 Chart
```{r urc8 numbers chart}
plot_ly(shared_urc8_all_ch, 
        x = ~age_group, y = ~individuals, 
        meta = ~ur8_name, split = ~ur8_name,
        hovertext = ~format(individuals, big.mark=',', scientific = FALSE), type = 'bar', 
        color = ~ur8_name, colors = unname(phsstyles::phs_colour_values)[1:8],
        hovertemplate = paste0("%{x} <br>", "%{meta} <br>", "Individuals: %{hovertext} <br>")) %>% 
  layout(clickmode = "none", showlegend = TRUE, margin = list(l = -5, b = 10, t = 40),
         xaxis = list(title = "Age Group", showline = FALSE, linecolor = "#b3b3b3",
                      categoryorder = "trace", tickangle = -45, titlefont = list(size = 16)),
         yaxis = list(title = "Count", linecolor = "#b3b3b3", titlefont = list(size = 16))
  ) %>%
  config(displayModeBar = TRUE, doubleClick = F,
         modeBarButtonsToRemove = list('select2d', 'lasso2d', 'zoomIn2d', 
                                       'zoomOut2d', 'autoScale2d', 
                                       'toggleSpikelines', 
                                       'hoverCompareCartesian', 
                                       'hoverClosestCartesian', 'toImage'), 
         displaylogo = F, editable = F)
```

### UR8 table
```{r urc8 numbers table}
datatable(shared_urc8_all_table, 
          options = list(dom = 't',
                         pageLength = 42,
                         columnDefs = list(list(visible=FALSE, targets=c(11)))),
          rownames = F,
          fillContainer = T,
          selection = 'none',
          colnames = c("Year", "Area", "UR8", age_order, "key")) %>%
    formatCurrency(age_order, currency = "", interval = 3, mark = ",", digits = 0)
```

### UR6 Chart
```{r urc6 numbers chart}
plot_ly(shared_urc6_all_ch, 
        x = ~age_group, y = ~individuals, 
        meta = ~ur6_name, split = ~ur6_name,
        hovertext = ~format(individuals, big.mark=',', scientific = FALSE), type = 'bar', 
        color = ~ur6_name, colors = unname(phsstyles::phs_colour_values)[1:6],
        hovertemplate = paste0("%{x} <br>", "%{meta} <br>", "Individuals: %{hovertext} <br>")) %>% 
  layout(clickmode = "none", showlegend = TRUE, margin = list(l = -5, b = 10, t = 40),
         xaxis = list(title = "Age Group", showline = FALSE, linecolor = "#b3b3b3",
                      categoryorder = "trace", tickangle = -45, titlefont = list(size = 16)),
         yaxis = list(title = "Count", linecolor = "#b3b3b3", titlefont = list(size = 16))
  ) %>%
  config(displayModeBar = TRUE, doubleClick = F,
         modeBarButtonsToRemove = list('select2d', 'lasso2d', 'zoomIn2d', 
                                       'zoomOut2d', 'autoScale2d', 
                                       'toggleSpikelines', 
                                       'hoverCompareCartesian', 
                                       'hoverClosestCartesian', 'toImage'), 
         displaylogo = F, editable = F)
```

### UR6 table
```{r urc6 numbers table}
datatable(shared_urc6_all_table, 
          options = list(dom = 't',
                         pageLength = 42,
                         columnDefs = list(list(visible=FALSE, targets=c(11)))),
          rownames = F,
          fillContainer = T,
          selection = 'none',
          colnames = c("Year", "Area", "UR6", age_order, "key")) %>%
    formatCurrency(age_order, currency = "", interval = 3, mark = ",", digits = 0)
```

Row {.tabset}
-----------------------------------------------------------------------

### Dementia per 100,000 population UR8 Age Group chart
```{r urc8 rate chart}
plot_ly(shared_urc8_all_rate_ch,
        x = ~age_group, y = ~rate,
        meta = ~ur8_name, split = ~ur8_name,
        hovertext = ~format(round_half_up(rate, digits = 1), big.mark=',', scientific = FALSE), type = 'bar',
        color = ~ur8_name, colors = unname(phsstyles::phs_colour_values)[1:8],
        hovertemplate = paste0("%{x} <br>", "%{meta} <br>", "Rate per 100,000: %{hovertext} <br>")) %>% 
  layout(clickmode = "none", showlegend = TRUE, margin = list(l = -5, b = 10, t = 40),
         xaxis = list(title = "Age Group", showline = FALSE, linecolor = "#b3b3b3",
                      categoryorder = "trace", tickangle = -45, titlefont = list(size = 16)),
         yaxis = list(title = "Rate per 100,000", linecolor = "#b3b3b3", titlefont = list(size = 16))
  ) %>%
  config(displayModeBar = TRUE, doubleClick = F,
         modeBarButtonsToRemove = list('select2d', 'lasso2d', 'zoomIn2d', 
                                       'zoomOut2d', 'autoScale2d', 
                                       'toggleSpikelines', 
                                       'hoverCompareCartesian', 
                                       'hoverClosestCartesian', 'toImage'), 
         displaylogo = F, editable = F)
```

### Dementia per 100,000 population UR8 Age Group table
```{r urc8 rate table}
datatable(shared_urc8_all_rate_table, 
          options = list(dom = 't',
                         pageLength = 42,
                         columnDefs = list(list(visible=FALSE, targets=c(11)))),
            rownames = F,
            fillContainer = T,
            selection = 'none',
          colnames = c("Year", "Area", "UR8 Name", age_order, "key")) %>%
    formatCurrency(age_order, currency = "", interval = 3, mark = ",", digits = 1)
```

### Dementia per 100,000 population UR6 Age Group chart
```{r urc6 rate chart}
plot_ly(shared_urc6_all_rate_ch,
        x = ~age_group, y = ~rate,
        meta = ~ur6_name, split = ~ur6_name,
        hovertext = ~format(round_half_up(rate, digits = 1), big.mark=',', scientific = FALSE), type = 'bar',
        color = ~ur6_name, colors = unname(phsstyles::phs_colour_values)[1:6],
        hovertemplate = paste0("%{x} <br>", "%{meta} <br>", "Rate per 100,000: %{hovertext} <br>")) %>% 
  layout(clickmode = "none", showlegend = TRUE, margin = list(l = -5, b = 10, t = 40),
         xaxis = list(title = "Age Group", showline = FALSE, linecolor = "#b3b3b3",
                      categoryorder = "trace", tickangle = -45, titlefont = list(size = 16)),
         yaxis = list(title = "%", linecolor = "#b3b3b3", titlefont = list(size = 16))
  ) %>%
  config(displayModeBar = TRUE, doubleClick = F,
         modeBarButtonsToRemove = list('select2d', 'lasso2d', 'zoomIn2d', 
                                       'zoomOut2d', 'autoScale2d', 
                                       'toggleSpikelines', 
                                       'hoverCompareCartesian', 
                                       'hoverClosestCartesian', 'toImage'), 
         displaylogo = F, editable = F)
```

### Dementia per 100,000 population UR6 Age Group table
```{r urc6 rate table}
datatable(shared_urc6_all_rate_table, 
          options = list(dom = 't',
                         pageLength = 42,
                         columnDefs = list(list(visible=FALSE, targets=c(11)))),
            rownames = F,
            fillContainer = T,
            selection = 'none',
          colnames = c("Year", "Area", "UR6 Name", age_order, "key")) %>%
    formatCurrency(age_order, currency = "", interval = 3, mark = ",", digits = 1)
```

```{js}
function filter_default() {

  document.getElementById("area_select1").getElementsByClassName("selectized")
  [0].selectize.removeOption("");
  
  document.getElementById("area_select1").getElementsByClassName("selectized") 
  [0].selectize.setValue("Scotland", false);
  
  document.getElementById("area_select2").getElementsByClassName("selectized")
  [0].selectize.removeOption("");
  
  document.getElementById("area_select2").getElementsByClassName("selectized") 
  [0].selectize.setValue("Scotland", false);
  
  document.getElementById("area_select_deaths").getElementsByClassName("selectized")
  [0].selectize.removeOption("");
  
  document.getElementById("area_select_deaths").getElementsByClassName("selectized") 
  [0].selectize.setValue("Scotland", false);
  
  document.getElementById("area_select_simd").getElementsByClassName("selectized")
  [0].selectize.removeOption("");
  
  document.getElementById("area_select_simd").getElementsByClassName("selectized") 
  [0].selectize.setValue("Scotland", false);
  
  document.getElementById("area_select_simd2").getElementsByClassName("selectized")
  [0].selectize.removeOption("");
  
  document.getElementById("area_select_simd2").getElementsByClassName("selectized") 
  [0].selectize.setValue("Scotland", false);
  
  document.getElementById("year_select_simd2").getElementsByClassName("selectized")
  [0].selectize.removeOption("");
  
  document.getElementById("year_select_simd2").getElementsByClassName("selectized") 
  [0].selectize.setValue("2023/24", false);
  
  document.getElementById("area_select_urc").getElementsByClassName("selectized")
  [0].selectize.removeOption("");
  
  document.getElementById("area_select_urc").getElementsByClassName("selectized") 
  [0].selectize.setValue("Scotland", false);
  
  document.getElementById("area_select_urc2").getElementsByClassName("selectized")
  [0].selectize.removeOption("");
  
  document.getElementById("area_select_urc2").getElementsByClassName("selectized") 
  [0].selectize.setValue("Scotland", false);
  
  document.getElementById("year_select_urc").getElementsByClassName("selectized")
  [0].selectize.removeOption("");
  
  document.getElementById("year_select_urc").getElementsByClassName("selectized") 
  [0].selectize.setValue("2023/24", false);
  
  }
  
  $(document).ready(filter_default);

```