---
title: "Dementia Analysis Dashboard"

output:
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
    # smart: false
    # theme:
    # logo: ps_logo.png

date:  '`r Sys.Date()`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Dementia Prevalence

Inputs {.sidebar}
-----------------------------------------------------------------------

```{r filter_section}
# prev_filter_select
filter_select("area_select1", "Select an Area",
                                    shared_prev_65plus_chart, group = ~area, allLevels = T, multiple = F)
```

Row
-----------------------------------------------------------------------

### Dementia Prevalence - Aged 65+
```{r 65+ table}
datatable(shared_prev_65plus_table,
          options = list(dom = 't'),
          rownames = F,
          selection = 'none') %>%
  formatCurrency(c('female', 'male', 'total'),currency = "", interval = 3, mark = ",", digits = 0)
```


### Dementia Prevalence in HSCP - Aged 65+
```{r 65+ chart}
plot_ly(shared_prev_65plus_chart, x = ~year, y = ~individuals, meta = ~gender, hovertext = ~individuals, type = 'bar', 
          color = ~gender, colors = c("#9B4393", "#0078D4"),
          hovertemplate = paste0("%{x} <br>", "%{meta} <br>", "Individuals: %{hovertext} <br>")) %>% 
    layout(clickmode = "none", showlegend = TRUE, margin = list(l = -5, b = 10, t = 40),
           xaxis = list(title = "Integration Authority Area", showline = FALSE, linecolor = "#b3b3b3",
                        categoryorder = "trace", tickangle = -45, titlefont = list(size = 16)),
           yaxis = list(title = "Prevalence", linecolor = "#b3b3b3", titlefont = list(size = 16))
    ) %>%
    config(displayModeBar = TRUE, doubleClick = F,
           modeBarButtonsToRemove = list('select2d', 'lasso2d', 'zoomIn2d', 
                                         'zoomOut2d', 'autoScale2d', 
                                         'toggleSpikelines', 
                                         'hoverCompareCartesian', 
                                         'hoverClosestCartesian', 'toImage'), 
           displaylogo = F, editable = F)
```

Row {.tabset}
-----------------------------------------------------------------------

### Dementia Rate per 100,000 - Aged 65+ Chart 
```{r rate chart}
plot_ly(shared_prev_65plus_chart, x = ~year, y = ~rate_100000, meta = ~gender, hovertext = ~individuals, 
        type = 'scatter', mode = 'lines', color = ~gender, colors = c("#9B4393", "#0078D4"),
        hovertemplate = paste0("%{x} <br>", "%{meta} <br>", "Individuals: %{hovertext} <br>")) %>% 
  layout(clickmode = "none", showlegend = TRUE, margin = list(l = -5, b = 10, t = 40),
         xaxis = list(title = "Financial Year", showline = FALSE, linecolor = "#b3b3b3",
                      categoryorder = "trace", tickangle = -45, titlefont = list(size = 16)),
         yaxis = list(title = "Prevalence rate per 100,000 pop (Aged 65+)", linecolor = "#b3b3b3", titlefont = list(size = 16), rangemode = "tozero")
  ) %>%
  config(displayModeBar = TRUE, doubleClick = F,
         modeBarButtonsToRemove = list('select2d', 'lasso2d', 'zoomIn2d', 
                                       'zoomOut2d', 'autoScale2d', 
                                       'toggleSpikelines', 
                                       'hoverCompareCartesian', 
                                       'hoverClosestCartesian', 'toImage'), 
         displaylogo = F, editable = F)
```

### Dementia Rate per 100,000 - Aged 65+ Table
```{r rate table}
datatable(shared_prev_65plus_rate_table, options = list(dom = 't'),
          rownames = F,
          selection = 'none') %>%
  formatCurrency(fy_dates, currency = "", interval = 3, mark = ",", digits = 0)
```


Row {.tabset}
-----------------------------------------------------------------------

### Dementia Age Group Split - Total
```{r age totals}
datatable(shared_prev_all_total, options = list(dom = 't'),
          rownames = F,
          selection = 'none') %>%
  formatCurrency(c(age_order, 'Total'), currency = "", interval = 3, mark = ",", digits = 0)
```

### Dementia Age Group Split - per 100,000 population
```{r age totals rate}
datatable(shared_prev_all_rate, options = list(dom = 't'),
          rownames = F,
          selection = 'none') %>%
  formatCurrency(age_order, currency = "", interval = 3, mark = ",", digits = 0)
```

Row
-----------------------------------------------------------------------

### Dementia Service Contacts & Patient Total 
```{r totals by source}
datatable(shared_prev_source, options = list(dom = 't'),
          rownames = F,
          selection = 'none') %>%
  formatCurrency(colnames(prev_all_source)[3:10], currency = "", interval = 3, mark = ",", digits = 0)
```

# Deaths

Inputs {.sidebar}
-----------------------------------------------------------------------

```{r filter_section deaths with}
# deaths_filter_select
filter_select("area_select_deaths", "Select an Area",
                                    shared_all_deaths_with_table, group = ~area, allLevels = T, multiple = F)
```

The figures included in this tab show dementia and non-dementia deaths. The figures show the proportion of deaths of those aged 65+ that were partially attributed to by dementia.

Row
-----------------------------------------------------------------------

### Deaths of those aged 65+
```{r all deaths with 65+}
datatable(shared_all_deaths_with_table, options = list(dom = 't'),
          rownames = F,
          selection = 'none') %>%
  formatCurrency(c('dementia', 'non_dementia'), currency = "", interval = 3, mark = ",", digits = 0)# %>% formatPercentage('proportion_to_dementia')
```

Row {.tabset}
-----------------------------------------------------------------------

### Proportion of deaths to that were attributed to dementia - age group (65+ only)
```{r proportion with chart age}
plot_ly(shared_deaths_with_chart_age, x = ~year, y = ~proportion, split = ~rev(age_group), meta = ~age_group, hovertext = ~proportion, type = 'bar', 
        color = ~age_group, colors = unname(phsstyles::phs_colour_values)[1:6],
        hovertemplate = paste0("%{x} <br>", "Age Group: %{meta} <br>", "Proportion of deaths attributed to dementia: %{hovertext}% <br>")) %>% 
  layout(clickmode = "none", showlegend = TRUE, margin = list(l = -5, b = 10, t = 40),
         xaxis = list(title = "Year", showline = FALSE, linecolor = "#b3b3b3",
                      categoryorder = "trace", tickangle = -45, titlefont = list(size = 16)),
         yaxis = list(title = "%", linecolor = "#b3b3b3", titlefont = list(size = 16))
  ) %>%
  config(displayModeBar = TRUE, doubleClick = F,
         modeBarButtonsToRemove = list('select2d', 'lasso2d', 'zoomIn2d', 
                                       'zoomOut2d', 'autoScale2d', 
                                       'toggleSpikelines', 
                                       'hoverCompareCartesian', 
                                       'hoverClosestCartesian', 'toImage'), 
         displaylogo = F, editable = F)
```

### Table
```{r proportion with table age}
datatable(shared_deaths_with_chart_age, options = list(dom = 't'),
          rownames = F,
          selection = 'none') %>%
  formatCurrency(c('dementia', 'non_dementia'), currency = "", interval = 3, mark = ",", digits = 0)
```

Row {.tabset}
-----------------------------------------------------------------------

### Proportion of deaths to to those that were attributed to dementia - gender (65+ only)
```{r proportion with chart gender}
plot_ly(shared_deaths_with_chart_gender, x = ~year, y = ~proportion, split = ~gender, meta = ~gender, hovertext = ~proportion, type = 'bar', 
        color = ~gender, colors = c("#9B4393", "#0078D4"),
        hovertemplate = paste0("%{x} <br>", "Gender: %{meta} <br>", "Proportion of deaths attributed to dementia: %{hovertext}% <br>")) %>% 
  layout(clickmode = "none", showlegend = TRUE, margin = list(l = -5, b = 10, t = 40),
         xaxis = list(title = "Year", showline = FALSE, linecolor = "#b3b3b3",
                      categoryorder = "trace", tickangle = -45, titlefont = list(size = 16)),
         yaxis = list(title = "%", linecolor = "#b3b3b3", titlefont = list(size = 16))
  ) %>%
  config(displayModeBar = TRUE, doubleClick = F,
         modeBarButtonsToRemove = list('select2d', 'lasso2d', 'zoomIn2d', 
                                       'zoomOut2d', 'autoScale2d', 
                                       'toggleSpikelines', 
                                       'hoverCompareCartesian', 
                                       'hoverClosestCartesian', 'toImage'), 
         displaylogo = F, editable = F)
```

### Table
```{r proportion with table gender}
datatable(shared_deaths_with_chart_gender, options = list(dom = 't'),
          rownames = F,
          selection = 'none') %>%
  formatCurrency(c('dementia', 'non_dementia'), currency = "", interval = 3, mark = ",", digits = 0)
```

# GP Dashboard Comparison

Inputs {.sidebar}
-----------------------------------------------------------------------

```{r filter_section GP comparison}
# deaths_filter_select
filter_select("area_select2", "Select an Area",
                                    shared_prev_comparison_65plus_table, group = ~area, allLevels = T, multiple = F)
```

The figures included in this tab prevalence from the dementia index compared to those from the [GP Disease Prevalence Dashboard](https://www.publichealthscotland.scot/publications/general-practice-disease-prevalence-data-visualisation/general-practice-disease-prevalence-visualisation-8-july-2025/).

Row
-----------------------------------------------------------------------

### Aged 65+ Comparison
```{r 65 plus comparison}
datatable(shared_prev_comparison_65plus_table, options = list(dom = 't'),
          rownames = F,
          selection = 'none') %>%
  formatCurrency(c('Female', 'Male', 'Total'), currency = "", interval = 3, mark = ",", digits = 0)
```

### Dementia Prevalence in HSCP 65+
```{r 65 plus chart comparison}
plot_ly(shared_prev_comparison_65plus_chart, 
        x = ~year, y = ~individuals, meta = ~group, hovertext = ~individuals, type = 'bar', 
        color = ~group, split = ~source, colors = unname(phsstyles::phs_colour_values)[1:4],
        hovertemplate = paste0("%{x} <br>", "%{meta} <br>", "Individuals: %{hovertext} <br>")) %>% 
  layout(clickmode = "none", showlegend = TRUE, margin = list(l = -5, b = 10, t = 40),
         xaxis = list(title = "Integration Authority Area", showline = FALSE, linecolor = "#b3b3b3",
                      categoryorder = "trace", tickangle = -45, titlefont = list(size = 16)),
         yaxis = list(title = "Prevalence", linecolor = "#b3b3b3", titlefont = list(size = 16))
  ) %>%
  config(displayModeBar = TRUE, doubleClick = F,
         modeBarButtonsToRemove = list('select2d', 'lasso2d', 'zoomIn2d', 
                                       'zoomOut2d', 'autoScale2d', 
                                       'toggleSpikelines', 
                                       'hoverCompareCartesian', 
                                       'hoverClosestCartesian', 'toImage'), 
         displaylogo = F, editable = F)
```

Row {.tabset}
-----------------------------------------------------------------------

### Age Group Split Comparisons
```{r age group comparison}
datatable(shared_prev_comparsion_ages_source, options = list(dom = 't'),
          rownames = F,
          selection = 'none') %>%
  formatCurrency(c(age_order2, 'Total'), currency = "", interval = 3, mark = ",", digits = 0)
```

# SIMD

Inputs {.sidebar}
-----------------------------------------------------------------------

**SIMD 65+ Age Group**

```{r filter_section SIMD}
filter_select("area_select_simd", "Select an Area",
                                    shared_simd_rates_65plus, group = ~area, allLevels = T, multiple = F)
```

***

<ul>
  <li>[**65+ Age Group**](#simd)</li>
  <li>[All Age Groups](#simd-all)</li>
</ul>

***

Row {.tabset}
-----------------------------------------------------------------------

### 65+ SIMD Chart
```{r 65+ simd numbers chart}
plot_ly(shared_simd_total_65plus_ch, 
        x = ~year, y = ~individuals, 
        meta = ~simd2020v2_sc_quintile, split = ~simd2020v2_sc_quintile,
        hovertext = ~individuals, type = 'bar', 
        color = ~simd2020v2_sc_quintile, colors = unname(phsstyles::phs_colour_values)[1:5],
        hovertemplate = paste0("%{x} <br>", "%{meta} <br>", "Individuals: %{hovertext} <br>")) %>% 
  layout(clickmode = "none", showlegend = TRUE, margin = list(l = -5, b = 10, t = 40),
         xaxis = list(title = "Integration Authority Area", showline = FALSE, linecolor = "#b3b3b3",
                      categoryorder = "trace", tickangle = -45, titlefont = list(size = 16)),
         yaxis = list(title = "Prevalence", linecolor = "#b3b3b3", titlefont = list(size = 16))
  ) %>%
  config(displayModeBar = TRUE, doubleClick = F,
         modeBarButtonsToRemove = list('select2d', 'lasso2d', 'zoomIn2d', 
                                       'zoomOut2d', 'autoScale2d', 
                                       'toggleSpikelines', 
                                       'hoverCompareCartesian', 
                                       'hoverClosestCartesian', 'toImage'), 
         displaylogo = F, editable = F)
```

### 65+ SIMD table
```{r 65+ simd numbers table}
datatable(shared_simd_total_65plus,# extensions = 'Scroller', 
          options = list(dom = 't',
                         pageLength = 40,
                         columnDefs = list(list(visible=FALSE, targets=c(7)))),
            rownames = F,
            fillContainer = T,
            selection = 'none') %>%
    formatCurrency(simd_65_cols, currency = "", interval = 3, mark = ",", digits = 0)
```

Row {.tabset}
-----------------------------------------------------------------------

### Dementia - per 100,000 population 65+ chart
```{r all age simd totals}
plot_ly(shared_simd_rates_65plus_ch, 
        x = ~year, y = ~rate, 
        meta = ~simd2020v2_sc_quintile, split = ~simd2020v2_sc_quintile,
        hovertext = ~rate, type = 'bar', 
        color = ~simd2020v2_sc_quintile, colors = unname(phsstyles::phs_colour_values)[1:5],
        hovertemplate = paste0("%{x} <br>", "%{meta} <br>", "Individuals: %{hovertext} <br>")) %>% 
  layout(clickmode = "none", showlegend = TRUE, margin = list(l = -5, b = 10, t = 40),
         xaxis = list(title = "Integration Authority Area", showline = FALSE, linecolor = "#b3b3b3",
                      categoryorder = "trace", tickangle = -45, titlefont = list(size = 16)),
         yaxis = list(title = "Prevalence", linecolor = "#b3b3b3", titlefont = list(size = 16))
  ) %>%
  config(displayModeBar = TRUE, doubleClick = F,
         modeBarButtonsToRemove = list('select2d', 'lasso2d', 'zoomIn2d', 
                                       'zoomOut2d', 'autoScale2d', 
                                       'toggleSpikelines', 
                                       'hoverCompareCartesian', 
                                       'hoverClosestCartesian', 'toImage'), 
         displaylogo = F, editable = F)
```


### Dementia - per 100,000 population 65+ table
```{r 65+ simd rates chart}
datatable(shared_simd_rates_65plus, options = list(dom = 't',
                                                   columnDefs = list(list(visible=FALSE, targets=c(7)))),
            rownames = F,
            fillContainer = TRUE,
            selection = 'none') %>%
    formatCurrency(simd_65_cols, currency = "", interval = 3, mark = ",", digits = 0)

```

# SIMD All {.hidden}

Inputs {.sidebar}
-----------------------------------------------------------------------

**SIMD All Age Groups**

```{r filter_section SIMD hidden}
filter_select("area_select_simd2", "Select an Area",
                                    shared_simd_rates_ch, group = ~area, allLevels = T, multiple = F)

filter_select("year_select_simd2", "Select a Financial Year",
                                    shared_simd_rates_ch, group = ~year, allLevels = T, multiple = F)
```

***

<ul>
  <li>[65+ Age Group](#simd)</li>
  <li>[**All Age Groups**](#simd-all)</li>
</ul>

***

Row {.tabset}
-----------------------------------------------------------------------

### SIMD Chart
```{r 65+ simd numbers chart hidden}
plot_ly(shared_simd_total_ch, 
        x = ~age_group, y = ~individuals, 
        meta = ~simd2020v2_sc_quintile, split = ~simd2020v2_sc_quintile,
        hovertext = ~individuals, type = 'bar', 
        color = ~simd2020v2_sc_quintile, colors = unname(phsstyles::phs_colour_values)[1:5],
        hovertemplate = paste0("%{x} <br>", "%{meta} <br>", "Individuals: %{hovertext} <br>")) %>% 
  layout(clickmode = "none", showlegend = TRUE, margin = list(l = -5, b = 10, t = 40),
         xaxis = list(title = "Age Group", showline = FALSE, linecolor = "#b3b3b3",
                      categoryorder = "trace", tickangle = -45, titlefont = list(size = 16)),
         yaxis = list(title = "Prevalence", linecolor = "#b3b3b3", titlefont = list(size = 16))
  ) %>%
  config(displayModeBar = TRUE, doubleClick = F,
         modeBarButtonsToRemove = list('select2d', 'lasso2d', 'zoomIn2d', 
                                       'zoomOut2d', 'autoScale2d', 
                                       'toggleSpikelines', 
                                       'hoverCompareCartesian', 
                                       'hoverClosestCartesian', 'toImage'), 
         displaylogo = F, editable = F)
```

### SIMD table
```{r 65+ simd numbers table hidden}
datatable(shared_simd_total,# extensions = 'Scroller', 
          options = list(dom = 't',
                         pageLength = 40,
                         columnDefs = list(list(visible=FALSE, targets=c(11)))),
            rownames = F,
            fillContainer = T,
            selection = 'none') %>%
    formatCurrency(age_order, currency = "", interval = 3, mark = ",", digits = 0)
```

Row {.tabset}
-----------------------------------------------------------------------

### Dementia Age Group SIMD Split - per 100,000 population chart
```{r all age simd totals hidden}
plot_ly(shared_simd_rates_ch,
        x = ~age_group, y = ~rate,
        meta = ~simd2020v2_sc_quintile, split = ~simd2020v2_sc_quintile,
        hovertext = ~rate, type = 'bar',
        color = ~simd2020v2_sc_quintile, colors = unname(phsstyles::phs_colour_values)[1:5],
        hovertemplate = paste0("%{x} <br>", "%{meta} <br>", "Individuals: %{hovertext} <br>")) %>%
  layout(clickmode = "none", showlegend = TRUE, margin = list(l = -5, b = 10, t = 40),
         xaxis = list(title = "Age Group", showline = FALSE, linecolor = "#b3b3b3",
                      categoryorder = "trace", tickangle = -45, titlefont = list(size = 16)),
         yaxis = list(title = "Prevalence", linecolor = "#b3b3b3", titlefont = list(size = 16))
  ) %>%
  config(displayModeBar = TRUE, doubleClick = F,
         modeBarButtonsToRemove = list('select2d', 'lasso2d', 'zoomIn2d',
                                       'zoomOut2d', 'autoScale2d',
                                       'toggleSpikelines',
                                       'hoverCompareCartesian',
                                       'hoverClosestCartesian', 'toImage'),
         displaylogo = F, editable = F)
```


### Dementia Age Group SIMD Split - per 100,000 population table
```{r simd rates table hidden}
datatable(shared_simd_rates, options = list(dom = 't',
                                            columnDefs = list(list(visible=FALSE, targets=c(11)))),
            rownames = F,
            fillContainer = TRUE,
            selection = 'none') %>%
    formatCurrency(age_order, currency = "", interval = 3, mark = ",", digits = 0)

```

# Urban Rural Classification

Inputs {.sidebar}
-----------------------------------------------------------------------

**URC 65+ Age Group**

```{r filter_section urc}
filter_select("area_select_urc", "Select an Area",
                                    shared_urc8_65plus_ch, group = ~area, allLevels = T, multiple = F)
```

***

<ul>
  <li>[**65+ Age Group**](#urban-rural-classification)</li>
  <li>[All Age Groups](#urban-rural-classification-all)</li>
</ul>

***

Row {.tabset}
-----------------------------------------------------------------------

### 65+ UR8 Chart
```{r 65+ urc8 numbers chart}
plot_ly(shared_urc8_65plus_ch, 
        x = ~year, y = ~individuals, 
        meta = ~ur8_name, split = ~ur8_name,
        hovertext = ~individuals, type = 'bar', 
        color = ~ur8_name, colors = unname(phsstyles::phs_colour_values)[1:8],
        hovertemplate = paste0("%{x} <br>", "%{meta} <br>", "Individuals: %{hovertext} <br>")) %>% 
  layout(clickmode = "none", showlegend = TRUE, margin = list(l = -5, b = 10, t = 40),
         xaxis = list(title = "Financial Year", showline = FALSE, linecolor = "#b3b3b3",
                      categoryorder = "trace", tickangle = -45, titlefont = list(size = 16)),
         yaxis = list(title = "Prevalence", linecolor = "#b3b3b3", titlefont = list(size = 16))
  ) %>%
  config(displayModeBar = TRUE, doubleClick = F,
         modeBarButtonsToRemove = list('select2d', 'lasso2d', 'zoomIn2d', 
                                       'zoomOut2d', 'autoScale2d', 
                                       'toggleSpikelines', 
                                       'hoverCompareCartesian', 
                                       'hoverClosestCartesian', 'toImage'), 
         displaylogo = F, editable = F)
```

### 65+ UR8 table
```{r 65+ urc8 numbers table}
datatable(shared_urc8_65plus_table, 
          options = list(dom = 't',
                         pageLength = 40,
                         columnDefs = list(list(visible=FALSE, targets=c(10)))),
            rownames = F,
            fillContainer = T,
            selection = 'none') %>%
    formatCurrency(urc8_cols, currency = "", interval = 3, mark = ",", digits = 0)
```


### 65+ UR6 Chart
```{r 65+ urc6 numbers chart}
plot_ly(shared_urc6_65plus_ch, 
        x = ~year, y = ~individuals, 
        meta = ~ur6_name, split = ~ur6_name,
        hovertext = ~individuals, type = 'bar', 
        color = ~ur6_name, colors = unname(phsstyles::phs_colour_values)[1:6],
        hovertemplate = paste0("%{x} <br>", "%{meta} <br>", "Individuals: %{hovertext} <br>")) %>% 
  layout(clickmode = "none", showlegend = TRUE, margin = list(l = -5, b = 10, t = 40),
         xaxis = list(title = "Financial Year", showline = FALSE, linecolor = "#b3b3b3",
                      categoryorder = "trace", tickangle = -45, titlefont = list(size = 16)),
         yaxis = list(title = "Prevalence", linecolor = "#b3b3b3", titlefont = list(size = 16))
  ) %>%
  config(displayModeBar = TRUE, doubleClick = F,
         modeBarButtonsToRemove = list('select2d', 'lasso2d', 'zoomIn2d', 
                                       'zoomOut2d', 'autoScale2d', 
                                       'toggleSpikelines', 
                                       'hoverCompareCartesian', 
                                       'hoverClosestCartesian', 'toImage'), 
         displaylogo = F, editable = F)
```

### 65+ UR6 table
```{r 65+ urc6 numbers table}
datatable(shared_urc6_65plus_table, 
          options = list(dom = 't',
                         pageLength = 40,
                         columnDefs = list(list(visible=FALSE, targets=c(8)))),
            rownames = F,
            fillContainer = T,
            selection = 'none') %>%
    formatCurrency(urc6_cols, currency = "", interval = 3, mark = ",", digits = 0)
```

Row {.tabset}
-----------------------------------------------------------------------

### Dementia UR8 65+ Percentage chart
```{r 65+ urc8 perc chart}
plot_ly(shared_urc8_65plus_ch, 
        x = ~year, y = ~perc*100, 
        meta = ~ur8_name, split = ~ur8_name,
        hovertext = ~perc*100, type = 'bar', 
        color = ~ur8_name, colors = unname(phsstyles::phs_colour_values)[1:8],
        hovertemplate = paste0("%{x} <br>", "%{meta} <br>", "Individuals: %{hovertext}% <br>")) %>% 
  layout(clickmode = "none", showlegend = TRUE, margin = list(l = -5, b = 10, t = 40),
         xaxis = list(title = "Financial Year", showline = FALSE, linecolor = "#b3b3b3",
                      categoryorder = "trace", tickangle = -45, titlefont = list(size = 16)),
         yaxis = list(title = "Prevalence", linecolor = "#b3b3b3", titlefont = list(size = 16))
  ) %>%
  config(displayModeBar = TRUE, doubleClick = F,
         modeBarButtonsToRemove = list('select2d', 'lasso2d', 'zoomIn2d', 
                                       'zoomOut2d', 'autoScale2d', 
                                       'toggleSpikelines', 
                                       'hoverCompareCartesian', 
                                       'hoverClosestCartesian', 'toImage'), 
         displaylogo = F, editable = F)
```


### Dementia UR8 65+ Percentage table
```{r 65+ urc8 perc table}
datatable(shared_urc8_65plus_perc_table, 
          options = list(dom = 't',
                         pageLength = 40,
                         columnDefs = list(list(visible=FALSE, targets=c(10)))),
            rownames = F,
            fillContainer = T,
            selection = 'none') %>%
    formatPercentage(urc8_cols, interval = 3, digits = 1)

```

### Dementia UR6 65+ Percentage chart
```{r 65+ urc6 perc chart}
plot_ly(shared_urc6_65plus_ch, 
        x = ~year, y = ~perc*100, 
        meta = ~ur6_name, split = ~ur6_name,
        hovertext = ~perc*100, type = 'bar', 
        color = ~ur6_name, colors = unname(phsstyles::phs_colour_values)[1:6],
        hovertemplate = paste0("%{x} <br>", "%{meta} <br>", "Individuals: %{hovertext}% <br>")) %>% 
  layout(clickmode = "none", showlegend = TRUE, margin = list(l = -5, b = 10, t = 40),
         xaxis = list(title = "Financial Year", showline = FALSE, linecolor = "#b3b3b3",
                      categoryorder = "trace", tickangle = -45, titlefont = list(size = 16)),
         yaxis = list(title = "Prevalence", linecolor = "#b3b3b3", titlefont = list(size = 16))
  ) %>%
  config(displayModeBar = TRUE, doubleClick = F,
         modeBarButtonsToRemove = list('select2d', 'lasso2d', 'zoomIn2d', 
                                       'zoomOut2d', 'autoScale2d', 
                                       'toggleSpikelines', 
                                       'hoverCompareCartesian', 
                                       'hoverClosestCartesian', 'toImage'), 
         displaylogo = F, editable = F)
```


### Dementia UR6 65+ Percentage table
```{r 65+ urc6 perc table}
datatable(shared_urc6_65plus_perc_table, 
          options = list(dom = 't',
                         pageLength = 40,
                         columnDefs = list(list(visible=FALSE, targets=c(8)))),
            rownames = F,
            fillContainer = T,
            selection = 'none') %>%
    formatPercentage(urc6_cols, interval = 3, digits = 1)

```

# Urban Rural Classification All {.hidden}

Inputs {.sidebar}
-----------------------------------------------------------------------

**URC All Age Groups**

```{r filter_section urc all age}
filter_select("area_select_urc2", "Select an Area",
                                    shared_urc8_all_ch, group = ~area, allLevels = T, multiple = F)
filter_select("year_select_urc", "Select a Financial Year",
                                    shared_urc8_all_ch, group = ~year, allLevels = T, multiple = F)
```

***

<ul>
  <li>[65+ Age Group](#urban-rural-classification)</li>
  <li>[**All Age Groups**](#urban-rural-classification-all)</li>
</ul>

***

Row {.tabset}
-----------------------------------------------------------------------

### UR8 Chart
```{r urc8 numbers chart}
plot_ly(shared_urc8_all_ch, 
        x = ~age_group, y = ~individuals, 
        meta = ~ur8_name, split = ~ur8_name,
        hovertext = ~individuals, type = 'bar', 
        color = ~ur8_name, colors = unname(phsstyles::phs_colour_values)[1:8],
        hovertemplate = paste0("%{x} <br>", "%{meta} <br>", "Individuals: %{hovertext} <br>")) %>% 
  layout(clickmode = "none", showlegend = TRUE, margin = list(l = -5, b = 10, t = 40),
         xaxis = list(title = "Age Group", showline = FALSE, linecolor = "#b3b3b3",
                      categoryorder = "trace", tickangle = -45, titlefont = list(size = 16)),
         yaxis = list(title = "Prevalence", linecolor = "#b3b3b3", titlefont = list(size = 16))
  ) %>%
  config(displayModeBar = TRUE, doubleClick = F,
         modeBarButtonsToRemove = list('select2d', 'lasso2d', 'zoomIn2d', 
                                       'zoomOut2d', 'autoScale2d', 
                                       'toggleSpikelines', 
                                       'hoverCompareCartesian', 
                                       'hoverClosestCartesian', 'toImage'), 
         displaylogo = F, editable = F)
```

### UR8 table
```{r urc8 numbers table}
datatable(shared_urc8_all_table, 
          options = list(dom = 't',
                         pageLength = 40,
                         columnDefs = list(list(visible=FALSE, targets=c(11)))),
            rownames = F,
            fillContainer = T,
            selection = 'none') %>%
    formatCurrency(age_order, currency = "", interval = 3, mark = ",", digits = 0)
```

### UR6 Chart
```{r urc6 numbers chart}
plot_ly(shared_urc6_all_ch, 
        x = ~age_group, y = ~individuals, 
        meta = ~ur6_name, split = ~ur6_name,
        hovertext = ~individuals, type = 'bar', 
        color = ~ur6_name, colors = unname(phsstyles::phs_colour_values)[1:6],
        hovertemplate = paste0("%{x} <br>", "%{meta} <br>", "Individuals: %{hovertext} <br>")) %>% 
  layout(clickmode = "none", showlegend = TRUE, margin = list(l = -5, b = 10, t = 40),
         xaxis = list(title = "Age Group", showline = FALSE, linecolor = "#b3b3b3",
                      categoryorder = "trace", tickangle = -45, titlefont = list(size = 16)),
         yaxis = list(title = "Prevalence", linecolor = "#b3b3b3", titlefont = list(size = 16))
  ) %>%
  config(displayModeBar = TRUE, doubleClick = F,
         modeBarButtonsToRemove = list('select2d', 'lasso2d', 'zoomIn2d', 
                                       'zoomOut2d', 'autoScale2d', 
                                       'toggleSpikelines', 
                                       'hoverCompareCartesian', 
                                       'hoverClosestCartesian', 'toImage'), 
         displaylogo = F, editable = F)
```

### UR6 table
```{r urc6 numbers table}
datatable(shared_urc6_all_table, 
          options = list(dom = 't',
                         pageLength = 40,
                         columnDefs = list(list(visible=FALSE, targets=c(11)))),
            rownames = F,
            fillContainer = T,
            selection = 'none') %>%
    formatCurrency(age_order, currency = "", interval = 3, mark = ",", digits = 0)
```

Row {.tabset}
-----------------------------------------------------------------------

### Dementia UR8 Age Group Split - Percentage chart
```{r urc8 perc chart}
plot_ly(shared_urc8_all_ch, 
        x = ~age_group, y = ~perc*100, 
        meta = ~ur8_name, split = ~ur8_name,
        hovertext = ~perc*100, type = 'bar', 
        color = ~ur8_name, colors = unname(phsstyles::phs_colour_values)[1:8],
        hovertemplate = paste0("%{x} <br>", "%{meta} <br>", "Individuals: %{hovertext}% <br>")) %>% 
  layout(clickmode = "none", showlegend = T, margin = list(l = -5, b = 10, t = 40),
         xaxis = list(title = "Age Group", showline = FALSE, linecolor = "#b3b3b3",
                      categoryorder = "trace", tickangle = -45, titlefont = list(size = 16)),
         yaxis = list(title = "Prevalence", linecolor = "#b3b3b3", titlefont = list(size = 16))
  ) %>%
  config(displayModeBar = TRUE, doubleClick = F,
         modeBarButtonsToRemove = list('select2d', 'lasso2d', 'zoomIn2d', 
                                       'zoomOut2d', 'autoScale2d', 
                                       'toggleSpikelines', 
                                       'hoverCompareCartesian', 
                                       'hoverClosestCartesian', 'toImage'), 
         displaylogo = F, editable = F)
```


### Dementia UR8 Age Group Split - Percentage table
```{r urc8 perc table}
datatable(shared_urc8_all_perc_table, 
          options = list(dom = 't',
                         pageLength = 40,
                         columnDefs = list(list(visible=FALSE, targets=c(11)))),
            rownames = F,
            fillContainer = T,
            selection = 'none') %>%
    formatPercentage(age_order, interval = 3, digits = 1)

```

### Dementia UR6 Age Group Split - Percentage chart
```{r urc6 perc chart}
plot_ly(shared_urc6_all_ch, 
        x = ~age_group, y = ~perc*100, 
        meta = ~ur6_name, split = ~ur6_name,
        hovertext = ~perc*100, type = 'bar', 
        color = ~ur6_name, colors = unname(phsstyles::phs_colour_values)[1:6],
        hovertemplate = paste0("%{x} <br>", "%{meta} <br>", "Individuals: %{hovertext}% <br>")) %>% 
  layout(clickmode = "none", showlegend = T, margin = list(l = -5, b = 10, t = 40),
         xaxis = list(title = "Age Group", showline = FALSE, linecolor = "#b3b3b3",
                      categoryorder = "trace", tickangle = -45, titlefont = list(size = 16)),
         yaxis = list(title = "Prevalence", linecolor = "#b3b3b3", titlefont = list(size = 16))
  ) %>%
  config(displayModeBar = TRUE, doubleClick = F,
         modeBarButtonsToRemove = list('select2d', 'lasso2d', 'zoomIn2d', 
                                       'zoomOut2d', 'autoScale2d', 
                                       'toggleSpikelines', 
                                       'hoverCompareCartesian', 
                                       'hoverClosestCartesian', 'toImage'), 
         displaylogo = F, editable = F)
```


### Dementia UR6 Age Group Split - Percentage table
```{r urc6 perc table}
datatable(shared_urc6_all_perc_table, 
          options = list(dom = 't',
                         pageLength = 40,
                         columnDefs = list(list(visible=FALSE, targets=c(11)))),
            rownames = F,
            fillContainer = T,
            selection = 'none') %>%
    formatPercentage(age_order, interval = 3, digits = 1)

```

```{js}
function filter_default() {

  document.getElementById("area_select1").getElementsByClassName("selectized")
  [0].selectize.removeOption("");
  
  document.getElementById("area_select1").getElementsByClassName("selectized") 
  [0].selectize.setValue("Scotland", false);
  
  document.getElementById("area_select2").getElementsByClassName("selectized")
  [0].selectize.removeOption("");
  
  document.getElementById("area_select2").getElementsByClassName("selectized") 
  [0].selectize.setValue("Scotland", false);
  
  document.getElementById("area_select_deaths").getElementsByClassName("selectized")
  [0].selectize.removeOption("");
  
  document.getElementById("area_select_deaths").getElementsByClassName("selectized") 
  [0].selectize.setValue("Scotland", false);
  
  document.getElementById("area_select_simd").getElementsByClassName("selectized")
  [0].selectize.removeOption("");
  
  document.getElementById("area_select_simd").getElementsByClassName("selectized") 
  [0].selectize.setValue("Scotland", false);
  
  document.getElementById("area_select_simd2").getElementsByClassName("selectized")
  [0].selectize.removeOption("");
  
  document.getElementById("area_select_simd2").getElementsByClassName("selectized") 
  [0].selectize.setValue("Scotland", false);
  
  document.getElementById("year_select_simd2").getElementsByClassName("selectized")
  [0].selectize.removeOption("");
  
  document.getElementById("year_select_simd2").getElementsByClassName("selectized") 
  [0].selectize.setValue("2023/24", false);
  
  document.getElementById("area_select_urc").getElementsByClassName("selectized")
  [0].selectize.removeOption("");
  
  document.getElementById("area_select_urc").getElementsByClassName("selectized") 
  [0].selectize.setValue("Scotland", false);
  
  document.getElementById("area_select_urc2").getElementsByClassName("selectized")
  [0].selectize.removeOption("");
  
  document.getElementById("area_select_urc2").getElementsByClassName("selectized") 
  [0].selectize.setValue("Scotland", false);
  
  document.getElementById("year_select_urc").getElementsByClassName("selectized")
  [0].selectize.removeOption("");
  
  document.getElementById("year_select_urc").getElementsByClassName("selectized") 
  [0].selectize.setValue("2023/24", false);
  
  }
  
  $(document).ready(filter_default);

```