---
title: "Dementia Analysis Dashboard"

output:
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
    # theme:
    # logo: ps_logo.png

date:  '`r Sys.Date()`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Dementia Prevalence

Inputs {.sidebar}
-----------------------------------------------------------------------

```{r filter_section}
# prev_filter_selec
filter_select("area_select1", "Select an Area",
                                    shared_prev_65plus_chart, group = ~area, allLevels = T, multiple = F)
```



Row
-----------------------------------------------------------------------

### Dementia Prevalence - Aged 65+
```{r 65+ table}
datatable(shared_prev_65plus_table,
          options = list(dom = 't'),
          rownames = F,
          selection = 'none')
```


### Dementia Prevalence in HSCP - Aged 65+
```{r 65+ chart}
plot_ly(shared_prev_65plus_chart, x = ~year, y = ~individuals, meta = ~gender, hovertext = ~individuals, type = 'bar', 
          color = ~gender, colors = c("#9B4393", "#0078D4"),
          hovertemplate = paste0("%{x} <br>", "%{meta} <br>", "Individuals: %{hovertext} <br>")) %>% 
    layout(clickmode = "none", showlegend = FALSE, margin = list(l = -5, b = 10, t = 40),
           xaxis = list(title = "Integration Authority Area", showline = FALSE, linecolor = "#b3b3b3",
                        categoryorder = "trace", tickangle = -45, titlefont = list(size = 16)),
           yaxis = list(title = "Prevalence", linecolor = "#b3b3b3", titlefont = list(size = 16))
    ) %>%
    config(displayModeBar = TRUE, doubleClick = F,
           modeBarButtonsToRemove = list('select2d', 'lasso2d', 'zoomIn2d', 
                                         'zoomOut2d', 'autoScale2d', 
                                         'toggleSpikelines', 
                                         'hoverCompareCartesian', 
                                         'hoverClosestCartesian', 'toImage'), 
           displaylogo = F, editable = F)
```

Row {.tabset}
-----------------------------------------------------------------------

### Dementia Rate per 100,000 - Aged 65+ Chart 
```{r rate chart}
plot_ly(shared_prev_65plus_chart, x = ~year, y = ~rate_100000, meta = ~gender, hovertext = ~individuals, 
        type = 'scatter', mode = 'lines', color = ~gender, colors = c("#9B4393", "#0078D4"),
        hovertemplate = paste0("%{x} <br>", "%{meta} <br>", "Individuals: %{hovertext} <br>")) %>% 
  layout(clickmode = "none", showlegend = FALSE, margin = list(l = -5, b = 10, t = 40),
         xaxis = list(title = "Financial Year", showline = FALSE, linecolor = "#b3b3b3",
                      categoryorder = "trace", tickangle = -45, titlefont = list(size = 16)),
         yaxis = list(title = "Prevalence rate per 100,000 pop (Aged 65+)", linecolor = "#b3b3b3", titlefont = list(size = 16))
  ) %>%
  config(displayModeBar = TRUE, doubleClick = F,
         modeBarButtonsToRemove = list('select2d', 'lasso2d', 'zoomIn2d', 
                                       'zoomOut2d', 'autoScale2d', 
                                       'toggleSpikelines', 
                                       'hoverCompareCartesian', 
                                       'hoverClosestCartesian', 'toImage'), 
         displaylogo = F, editable = F)
```

### Dementia Rate per 100,000 - Aged 65+ Table
```{r rate table}
datatable(shared_prev_65plus_rate_table, options = list(dom = 't'),
          rownames = F,
          selection = 'none')
```


Row {.tabset}
-----------------------------------------------------------------------

### Dementia Age Group Split - Total
```{r age totals}
datatable(shared_prev_all_total, options = list(dom = 't'),
          rownames = F,
          selection = 'none')
```

### Dementia Age Group Split - per 100,000 population
```{r age totals rate}
datatable(shared_prev_all_rate, options = list(dom = 't'),
          rownames = F,
          selection = 'none')
```

Row
-----------------------------------------------------------------------

### Dementia Service Contacts & Patient Total 
```{r totals by source}
datatable(shared_prev_source, options = list(dom = 't'),
          rownames = F,
          selection = 'none')
```

# Deaths

Inputs {.sidebar}
-----------------------------------------------------------------------

```{r filter_section deaths with}
# deaths_filter_select
filter_select("area_select2", "Select an Area",
                                    shared_all_deaths_with_table, group = ~area, allLevels = T, multiple = F)
```

The figures included in this tab show dementia and non-dementia deaths. The figures show the proportion of deaths of those aged 65+ that were partially attributed to by dementia.

Row
-----------------------------------------------------------------------

### Deaths of those aged 65+
```{r all deaths with 65+}
datatable(shared_all_deaths_with_table, options = list(dom = 't'),
          rownames = F,
          selection = 'none')
```

Row {.tabset}
-----------------------------------------------------------------------

### Proportion of deaths to that were attributed to dementia - age group (65+ only)
```{r proportion with chart age}
plot_ly(shared_deaths_with_chart_age, x = ~year, y = ~proportion, split = ~rev(age_group), meta = ~age_group, hovertext = ~proportion, type = 'bar', 
        color = ~age_group, #colors = c("#9B4393", "#0078D4"),
        hovertemplate = paste0("%{x} <br>", "Age Group: %{meta} <br>", "Proportion of deaths attributed to dementia: %{hovertext}% <br>")) %>% 
  layout(clickmode = "none", showlegend = FALSE, margin = list(l = -5, b = 10, t = 40),
         xaxis = list(title = "Year", showline = FALSE, linecolor = "#b3b3b3",
                      categoryorder = "trace", tickangle = -45, titlefont = list(size = 16)),
         yaxis = list(title = "%", linecolor = "#b3b3b3", titlefont = list(size = 16))
  ) %>%
  config(displayModeBar = TRUE, doubleClick = F,
         modeBarButtonsToRemove = list('select2d', 'lasso2d', 'zoomIn2d', 
                                       'zoomOut2d', 'autoScale2d', 
                                       'toggleSpikelines', 
                                       'hoverCompareCartesian', 
                                       'hoverClosestCartesian', 'toImage'), 
         displaylogo = F, editable = F)
```

### Table
```{r proportion with table age}
datatable(shared_deaths_with_chart_age, options = list(dom = 't'),
          rownames = F,
          selection = 'none')
```
Row {.tabset}
-----------------------------------------------------------------------

### Proportion of deaths to to those that were attributed to dementia - gender (65+ only)
```{r proportion with chart gender}
plot_ly(shared_deaths_with_chart_gender, x = ~year, y = ~proportion, split = ~gender, meta = ~gender, hovertext = ~proportion, type = 'bar', 
        color = ~gender, colors = c("#9B4393", "#0078D4"),
        hovertemplate = paste0("%{x} <br>", "Gender: %{meta} <br>", "Proportion of deaths attributed to dementia: %{hovertext}% <br>")) %>% 
  layout(clickmode = "none", showlegend = FALSE, margin = list(l = -5, b = 10, t = 40),
         xaxis = list(title = "Year", showline = FALSE, linecolor = "#b3b3b3",
                      categoryorder = "trace", tickangle = -45, titlefont = list(size = 16)),
         yaxis = list(title = "%", linecolor = "#b3b3b3", titlefont = list(size = 16))
  ) %>%
  config(displayModeBar = TRUE, doubleClick = F,
         modeBarButtonsToRemove = list('select2d', 'lasso2d', 'zoomIn2d', 
                                       'zoomOut2d', 'autoScale2d', 
                                       'toggleSpikelines', 
                                       'hoverCompareCartesian', 
                                       'hoverClosestCartesian', 'toImage'), 
         displaylogo = F, editable = F)
```

### Table
```{r proportion with table gender}
datatable(shared_deaths_with_chart_gender, options = list(dom = 't'),
          rownames = F,
          selection = 'none')
```

```{js }
function filter_default() {

  document.getElementById("area_select1").getElementsByClassName("selectized")
  [0].selectize.removeOption("");
  
  document.getElementById("area_select1").getElementsByClassName("selectized") 
  [0].selectize.setValue("Aberdeen City", false);
  
  document.getElementById("area_select2").getElementsByClassName("selectized")
  [0].selectize.removeOption("");
  
  document.getElementById("area_select2").getElementsByClassName("selectized") 
  [0].selectize.setValue("Aberdeen City", false);
  
  document.getElementById("area_select3").getElementsByClassName("selectized")
  [0].selectize.removeOption("");
  
  document.getElementById("area_select3").getElementsByClassName("selectized") 
  [0].selectize.setValue("Aberdeen City", false);
  
  }
  
  $(document).ready(filter_default);

```